---
title: "Clark_County_main.R"
author: "ravi"
date: "4/30/2023"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

directory_loc = "/Users/ravigoyal/Dropbox/Academic/Research/Projects/HIVcounty/"
county_demo_file_subloc = "Clark_County_data/clark_county_demographics.csv"
county_zip_file_subloc = "Clark_County_data/clark_county_by_zip.csv"

```


```{r library, echo=FALSE, message=FALSE, warning = FALSE}
library('tidyverse')
library('readxl')
library('kableExtra')
library('eeptools')
library('geepack')
library('caret')
library('tidycensus')
library('zipcodeR')
library('leaflet')
library('classInt')
library('RColorBrewer')

source("/Users/ravigoyal/Dropbox/Academic/Research/Projects/HIVcounty/HIVcounty/HIVcounty_func.R")
```

## Research question

Research Question 3: Identify predictors (including viral clustering) that are associated with high rates of stage 3-4 new diagnoses. 

## Create analytic dataset

We create an analytic dataset that includes individual and zip code level covariates. 

```{r dataset, echo=FALSE, message=FALSE, warning = FALSE}

county_zip_list = search_county("Clark", "NV") %>% pull(zipcode)

county_diagnosis.df  = read_county_data(directory_loc = directory_loc,
                             county_demo_file_subloc = county_demo_file_subloc,
                             county_zip_file_subloc = county_zip_file_subloc,
                             county_zip_list = county_zip_list
                             ) 

```     

Number of individuals after filtering to diagnosed in county: `r county_diagnosis.df %>% nrow()`

## Outcome generation

We dichotomize the HIV stage of an individual.

* HIV ONLY (HIV, STAGE 1 OR 2) 

VS

* HIV AND AIDS SIMULTANEOUSLY (HIV, STAGE 3)
* HIV AND LATER AIDS (HIV, STAGE 3)


```{r outcomes, echo=FALSE, message=FALSE, warning = FALSE}

county_diagnosis.df = county_diagnosis.df %>%
  mutate(outcome_disease_categ_dx_bin = case_when(
    `Disease Category` == "HIV ONLY (HIV, STAGE 1 OR 2)" ~ 0,
    `Disease Category` == "HIV AND AIDS SIMULTANEOUSLY (HIV, STAGE 3)" ~ 1,
    `Disease Category` == "HIV AND LATER AIDS (HIV, STAGE 3)" ~ 1,
    TRUE ~ NA_real_))

county_diagnosis.df = county_diagnosis.df %>%
  mutate(outcome = outcome_disease_categ_dx_bin) %>%
  filter(!is.na(outcome))

```

## Descriptive statistics

The number of individuals in Stages 1 or 2 (outcome = 0) vs. number of individuals in Stage 3

```{r desc_stats, echo=FALSE, message=FALSE, warning = FALSE}

county_diagnosis.df %>% 
  group_by(outcome) %>%
  summarize(outcome_total = n()) %>%
  kbl() %>%
  kable_paper("hover", full_width = F)
  
```

### Geographic map

```{r desc_map, echo=FALSE, message=FALSE, warning = FALSE}

zcta_data <- get_acs(
  geography = "zcta",
  variables = "B19013_001",
  zcta = county_zip_list, 
  geometry = TRUE
)

county_summary.df = county_diagnosis.df %>%
  group_by(rsd_zip_cd) %>%
  summarize(n_tot = n(),
            n_outcome = sum(outcome)) %>%
  ungroup() %>%
  mutate(per_outcome = n_outcome/n_tot) %>%
  mutate(rsd_zip_cd_cat = as.character(rsd_zip_cd))
  

zcta_level.df = left_join(zcta_data,
                          county_summary.df,
                          by = c("GEOID" = "rsd_zip_cd_cat"))

zcta_level.df = left_join(zcta_level.df,
                          zip_code_db %>% select(zipcode, lat, lng),
                          by = c("GEOID" = "zipcode")) %>%
  filter(!is.na(lat))

pal_fun <- colorQuantile("YlOrRd", NULL, n = 5)

p_popup <- paste0("<strong>Proportion Late Diagnosis: </strong>", zcta_level.df$per_outcome)
breaks_qt <- classIntervals(zcta_level.df$per_outcome, n = 7, style = "quantile")

leaflet(zcta_level.df) %>%
  addPolygons(
    stroke = FALSE, 
    fillColor = ~pal_fun(per_outcome),
    fillOpacity = 0.8, smoothFactor = 0.5,
    popup = p_popup) %>%
  addTiles()  %>%
  addLegend("bottomright", 
            colors = brewer.pal(7, "YlOrRd"), 
            labels = paste0("up to ", format(breaks_qt$brks[-1], digits = 2)),
            title =  'Proportion of Late Diagnoses') %>%
  addCircleMarkers(zcta_level.df$lng, zcta_level.df$lat, radius=zcta_level.df$n_tot/50, 
                   color = "blue", 
                   stroke = FALSE, fillOpacity = 0.5) 

```
### Geographic map

```{r desc_demo, echo=FALSE, message=FALSE, warning = FALSE}

race_prop.df = county_diagnosis.df %>%
  group_by(Race.full.factor) %>%
  summarize(total_n = n(),
            late_diag_n = sum(outcome)) %>%
  ungroup() %>%
  mutate(prop = late_diag_n/total_n)

p_race <-ggplot(race_prop.df, aes(x=Race.full.factor, y=prop, fill=Race.full.factor)) +
  geom_bar(stat="identity")+theme_minimal() +
  geom_text(aes(label=total_n), vjust=-1) +
  labs(title="",
       x="Race / Ethnicity", y = "Proportion Late Diagnosis") +
  ylim(0,1) +
  theme(axis.title.x = element_text(size = 16),
        axis.text.x = element_blank(),
        axis.title.y = element_text(size = 16)) +
  guides(fill=guide_legend(title="Race / Ethnicity"))+
  scale_fill_brewer(palette="Spectral")
p_race

exp_cat_prop.df = county_diagnosis.df %>%
  group_by(`Exposure Category.factor`) %>%
  summarize(total_n = n(),
            late_diag_n = sum(outcome)) %>%
  ungroup() %>%
  mutate(prop = late_diag_n/total_n)

p_exp_cat <-ggplot(exp_cat_prop.df, aes(x=`Exposure Category.factor`, y=prop, fill=`Exposure Category.factor`)) +
  geom_bar(stat="identity")+theme_minimal() +
  geom_text(aes(label=total_n), vjust=-1) +
  labs(title="",
       x="Exposure Category", y = "Proportion Late Diagnosis") +
  ylim(0,1) +
  theme(axis.title.x = element_text(size = 16),
        axis.text.x = element_blank(),
        axis.title.y = element_text(size = 16)) +
  guides(fill=guide_legend(title="Exposure Risk")) +
  scale_fill_brewer(palette="Spectral")
p_exp_cat


edu_cat_prop.df = county_diagnosis.df %>%
  group_by(Education.factor) %>%
  summarize(total_n = n(),
            late_diag_n = sum(outcome)) %>%
  ungroup() %>%
  mutate(prop = late_diag_n/total_n)

p_edu_cat <-ggplot(edu_cat_prop.df, aes(x=Education.factor, y=prop, fill=Education.factor)) +
  geom_bar(stat="identity")+theme_minimal() +
  geom_text(aes(label=total_n), vjust=-1) +
  labs(title="",
       x="Education Category", y = "Proportion Late Diagnosis") +
  ylim(0,1) +
  theme(axis.title.x = element_text(size = 16),
        axis.text.x = element_blank(),
        axis.title.y = element_text(size = 16)) +
  guides(fill=guide_legend(title="Education")) +
  scale_fill_brewer(palette="Spectral")
p_edu_cat


```

## Individual-level analyses

### Regression

Individual-level variables to consider: 
* predictor_diagnosis_year
* Age.diagnosis
* predictor_hispanic
* predictor_idu
* predictor_msm
* predictor_education
* predictor_birth_sex
* clustered_2022

First set of estimates and p-values are for univariate regression. The second set is for multivariate regression.

```{r predictors, echo=FALSE, message=FALSE, warning = FALSE}

county_diagnosis.df = county_diagnosis.df %>%
  mutate(predictor_education = case_when(
    Education == "<= 8TH GRADE" ~ 0,
    Education == "SOME SCHOOL, LEVEL UNKNOWN" ~ 0,
    Education == "SOME HIGH SCHOOL" ~ 0,
    Education == "HIGH SCHOOL GRAD" ~ 0,
    Education == "SOME COLLEGE" ~ 1,
    Education == "COLLEGE DEGREE" ~ 1,
    Education == "POST-GRADUATE WORK" ~ 1,
    TRUE ~ NA_real_))

county_diagnosis.df = county_diagnosis.df %>%
  mutate(predictor_hispanic = case_when(
    Race == "WHITE" ~ 0,
    Race == "OTHER" ~ 0,
    Race == "BLACK" ~ 0,
    Race == "HISPANIC, ANY RACE" ~ 1,
    TRUE ~ NA_real_))

county_diagnosis.df = county_diagnosis.df %>%
  mutate(predictor_msm = case_when(
    `Exposure Category` == "MSM" ~ 1,
    `Exposure Category` == "IDU" ~ 0,
    `Exposure Category` == "NO REPORTED RISK" ~ 0,
    `Exposure Category` == "MSM & IDU" ~ 1,
    `Exposure Category` == "HETEROSEXUAL CONTACT" ~ 0,
    `Exposure Category` == "OTHER" ~ 0,
    `Exposure Category` == "PERINATAL EXPOSURE" ~ 0,
    TRUE ~ NA_real_))

county_diagnosis.df = county_diagnosis.df %>%
  mutate(predictor_idu = case_when(
    `Exposure Category` == "MSM" ~ 0,
    `Exposure Category` == "IDU" ~ 1,
    `Exposure Category` == "NO REPORTED RISK" ~ 0,
    `Exposure Category` == "MSM & IDU" ~ 1,
    `Exposure Category` == "HETEROSEXUAL CONTACT" ~ 0,
    `Exposure Category` == "OTHER" ~ 0,
    `Exposure Category` == "PERINATAL EXPOSURE" ~ 0,
    TRUE ~ NA_real_))

county_diagnosis.df = county_diagnosis.df %>%
  mutate(predictor_diagnosis_year = `Diagnosis year`)

county_diagnosis.df = county_diagnosis.df %>%
  mutate(predictor_birth_sex = case_when(
    `Birth Sex` == "FEMALE" ~ 0,
    `Birth Sex` == "MALE" ~ 1,
    TRUE ~ NA_real_))

```

```{r exploratory_analysis, echo=FALSE, message=FALSE, warning = FALSE}

p_predictor_diagnosis_year <- ggplot(county_diagnosis.df, aes(x=predictor_diagnosis_year, y=outcome)) +
  geom_point(col = "blue") +
  geom_smooth(col = "red") +
  labs(title="Late Diagnosis by Diagnosis Year",
        x ="Diagnosis Year", y = "Late Diagnosis")

p_predictor_diagnosis_year

p_Age.diagnosis <- ggplot(county_diagnosis.df, aes(x=Age.diagnosis, y=outcome)) +
  geom_point() +
  geom_smooth() +
  labs(title="Late Diagnosis by Age at Diagnosis",
        x ="Age at Diagnosis", y = "Late Diagnosis")

p_Age.diagnosis

p_zip_income <- ggplot(county_diagnosis.df, aes(x=Median_Income_Total, y=outcome)) +
  geom_point() +
  geom_smooth() +
  labs(title="Late Diagnosis by Zip Code Level Income",
        x ="Zip Code Level Income", y = "Late Diagnosis")

p_zip_income

p_percent_clustered_2022 <- ggplot(county_diagnosis.df, aes(x=percent_clustered_2022, y=outcome)) +
  geom_point() +
  geom_smooth() +
  labs(title="Late Diagnosis by Zip Code Clustering",
        x ="Zip Code Level Clustering", y = "Late Diagnosis")

p_percent_clustered_2022

```

```{r Univariate_reg, echo=FALSE, message=FALSE, warning = FALSE}

uni_reg_res.df = tibble(
  variable = NULL,
  estimate_uni = NULL,
  p_val_uni = NULL
)

reg_variable_ind.vec = c(
                 "predictor_diagnosis_year",
                 "Age.diagnosis",
                 "predictor_hispanic",
                 "predictor_idu",
                 "predictor_msm",
                 "predictor_education",
                 "predictor_birth_sex",
                 "clustered_2022")

for (reg_variable_ind in reg_variable_ind.vec) {

  regression_formula <- paste0("outcome", " ~ ", reg_variable_ind)

  uni_reg = glm(as.formula(regression_formula), 
              family = binomial(),
              data = county_diagnosis.df)

  uni_reg_sum = summary(uni_reg)
  uni_reg_res_TEMP.df = tibble(
    variable = reg_variable_ind,
    estimate_uni = uni_reg_sum$coefficients[2,c(1)],
    p_val_uni = uni_reg_sum$coefficients[2,c(4)]
  )
  
  uni_reg_res.df = bind_rows(uni_reg_res.df, uni_reg_res_TEMP.df)
}

```

```{r multivariate_reg, echo=FALSE, message=FALSE, warning = FALSE}

regression_formula <- paste0("outcome", " ~ ", paste0(reg_variable_ind.vec, collapse = " + "))

multi_reg = glm(as.formula(regression_formula), 
            family = binomial(),
            data = county_diagnosis.df)

multi_reg_sum = summary(multi_reg)
multi_reg_res.df = tibble(
  variable = rownames(multi_reg_sum$coefficients)[c(2:nrow(multi_reg_sum$coefficients))],
  estimate_multi = multi_reg_sum$coefficients[2:nrow(multi_reg_sum$coefficients),c(1)],
  p_val_multi = multi_reg_sum$coefficients[2:nrow(multi_reg_sum$coefficients),c(4)]
)


full_join(uni_reg_res.df, multi_reg_res.df, by = "variable") %>%
  kbl() %>%
  kable_paper("hover", full_width = F)

```
### Machine Learning: Random Forest

```{r machine_learning, echo=FALSE, message=FALSE, warning = FALSE}

county_diagnosis_ML.df = county_diagnosis.df %>%
  select(c(outcome, reg_variable_ind.vec))

county_diagnosis_ML.df <- county_diagnosis_ML.df[complete.cases(county_diagnosis_ML.df),]

county_diagnosis_ML.df = county_diagnosis_ML.df %>%
  mutate(outcome = as.factor(outcome))

set.seed(107)
inTrain <- createDataPartition(
  y = county_diagnosis_ML.df$outcome, ## the outcome data are needed
  p = .75, ## The percentage of data in the training set
  list = FALSE
)

training <- county_diagnosis_ML.df[ inTrain,]
testing  <- county_diagnosis_ML.df[-inTrain,]

trControl <- trainControl(method = "cv",
    number = 10,
    search = "grid")

rf_model = train(outcome ~ .,
      data = training,
      method = "rf",
      trControl = trainControl(),
      tuneGrid = NULL)


```

#### Plot: Variable importance

```{r machine_learning_plot, echo=FALSE, message=FALSE, warning = FALSE}

rf_model_Imp <- varImp(rf_model, scale = FALSE)

plot(rf_model_Imp)
```

#### Table: Prediction assessment

```{r machine_learning_prediction, echo=FALSE, message=FALSE, warning = FALSE}

rf_test <- predict(rf_model, newdata = testing)
confusionMatrix(rf_test, as.factor(testing$outcome))
```

## Zip Code-level predictors

### Regression

zip code-level variables to consider: 

* Median income
* Percent in poverty
* Percent clustered in 2022

First set of estimates and p-values are for univariate regression. The second set is for multivariate regression.

```{r county_predictors, echo=FALSE, message=FALSE, warning = FALSE}

county_diagnosis.df = county_diagnosis.df %>%
  mutate(predictor_county_per_poverty = Poverty / Gender_All)

```


```{r county_univariate_reg, echo=FALSE, message=FALSE, warning = FALSE}
 
reg_variable_zip.vec = c(
                 "Median_Income_Total",
                 "predictor_county_per_poverty",
                 "percent_clustered_2022"
                 )

reg_variable.vec = c(reg_variable_ind.vec, reg_variable_zip.vec)

uni_reg_res.df = tibble(
  variable = NULL,
  estimate_uni = NULL,
  p_val_uni = NULL
)

for (reg_variable in reg_variable.vec) {

  regression_formula <- paste0("outcome", " ~ ", reg_variable)

  uni_reg = glm(as.formula(regression_formula), 
              family = binomial(),
              data = county_diagnosis.df)

  uni_reg_sum = summary(uni_reg)
  uni_reg_res_TEMP.df = tibble(
    variable = reg_variable,
    estimate_uni = uni_reg_sum$coefficients[2,c(1)],
    p_val_uni = uni_reg_sum$coefficients[2,c(4)]
  )
  
  uni_reg_res.df = bind_rows(uni_reg_res.df, uni_reg_res_TEMP.df)
}

```

```{r county_multivariate_reg, echo=FALSE, message=FALSE, warning = FALSE}

regression_formula <- paste0("outcome", " ~ ", paste0(reg_variable.vec, collapse = " + "))

multi_reg = glm(as.formula(regression_formula), 
            family = binomial(),
            data = county_diagnosis.df)

multi_reg_sum = summary(multi_reg)
multi_reg_res.df = tibble(
  variable = rownames(multi_reg_sum$coefficients)[c(2:nrow(multi_reg_sum$coefficients))],
  estimate_multi = multi_reg_sum$coefficients[2:nrow(multi_reg_sum$coefficients),c(1)],
  p_val_multi = multi_reg_sum$coefficients[2:nrow(multi_reg_sum$coefficients),c(4)]
)


full_join(uni_reg_res.df, multi_reg_res.df, by = "variable") %>%
  kbl() %>%
  kable_paper("hover", full_width = F)

```
### Machine_learning

```{r county_machine_learning, echo=FALSE, message=FALSE, warning = FALSE}

county_diagnosis_ML.df = county_diagnosis.df %>%
  select(c(outcome, reg_variable.vec))

county_diagnosis_ML.df <- county_diagnosis_ML.df[complete.cases(county_diagnosis_ML.df),]

county_diagnosis_ML.df = county_diagnosis_ML.df %>%
  mutate(outcome = as.factor(outcome))

set.seed(107)
inTrain <- createDataPartition(
  y = county_diagnosis_ML.df$outcome, ## the outcome data are needed
  p = .75, ## The percentage of data in the training set
  list = FALSE
)

training <- county_diagnosis_ML.df[ inTrain,]
testing  <- county_diagnosis_ML.df[-inTrain,]

trControl <- trainControl(method = "cv",
    number = 10,
    search = "grid")

rf_model_county = train(outcome ~ .,
      data = training,
      method = "rf",
      trControl = trainControl(),
      tuneGrid = NULL)


```

#### Plot: Variable importance

```{r county_machine_learning_plot, echo=FALSE, message=FALSE, warning = FALSE}

rf_model_Imp <- varImp(rf_model_county, scale = FALSE)

plot(rf_model_Imp)
```

#### Table: Prediction

```{r county_machine_learning_prediction, echo=FALSE, message=FALSE, warning = FALSE}

rf_test <- predict(rf_model_county, newdata = testing)
confusionMatrix(rf_test, as.factor(testing$outcome))
```
