---
title: "Clark County Continuum of Care analysis"
number-sections: true
# toc: true
toc-depth: 4
format: 
  html:
    page-layout: full
    self-contained: true
editor_options: 
  chunk_output_type: console
---

```{r}
#| label: load-packages
#| echo: false
#| include: false
options(tigris_use_cache = TRUE)

library('MASS')
library('tidyverse')
library('readxl')
library('kableExtra')
library('eeptools')
library('geepack')
library('caret')
library('tidycensus')
library('zipcodeR')
library('leaflet')
library('classInt')
library('RColorBrewer')
library('gtsummary')
library('lme4')
library('glmmLasso')
library('nlme')
library('mapview')
library('tidycensus')
library('here')
library('gt')
library('sf')
data("zip_code_db")
source("HIVcounty_func.R")


  



```

# Create analytic dataset

We create an analytic dataset that includes individual and zip code level covariates. 


```{r}
#| label: analytic-dataset
#| echo: false
#| include: false

# load in alltables
#use output till we update alltables file in Current Data
# onedrive_path <- readLines(here("OneDrivePath.txt"))
# # load(paste0(onedrive_path, "Current_Data/UCSD_eHARS_alltables_052023.Rdata"))))
#use output till we update alltables file in Current Data
#load("~/Documents/Projects/Nevada/data #requests/NV_Data_Processing_Scripts/output/Nevada_eHARS_alltables_073123.Rdata")

# directory_loc = "/Users/ravigoyal/Dropbox/Academic/Research/Projects/HIVcounty/Clark_County_data/"

# load(paste(directory_loc, "Nevada_eHARS_alltables_073123.Rdata", sep=""))

directory_loc = paste0(here(),'/')

result_file_loc = paste(directory_loc, "Results/", sep = "")

county_demo_file_subloc = "Nevada_eHARS_demographics_073123.csv"
county_zip_file_subloc =  "Nevada_eHARS_by_zip.csv_073123.csv"
county_test_file_subloc = "2023 SNHD testing sites.xlsx"

county_zip_list = search_county("Clark", "NV") %>% pull(zipcode)
county_zip_list_full = search_county("Clark", "NV")

testing_site_bool = TRUE


################
## Create analytic dataset
################
directory_loc = directory_loc
county_demo_file_subloc = county_demo_file_subloc
county_zip_file_subloc = county_zip_file_subloc

county.df  = read_county_data(directory_loc = directory_loc,
                              county_demo_file_subloc = county_demo_file_subloc,
                              county_zip_file_subloc = county_zip_file_subloc
                              # county_zip_list = county_zip_list
) 


```

# Download ACS data

        
* Population - B01003_001 
    + Estimate!!Total TOTAL POPULATION
* HISPANIC - B03001_003 
    + Estimate!!Total:!!Hispanic or Latino: HISPANIC OR LATINO ORIGIN BY SPECIFIC ORIGIN tract
* EDUCATION - B16010_015 
    + Estimate!!Total:!!High school graduate (includes equivalency): EDUCATIONAL ATTAINMENT AND EMPLOYMENT STATUS BY LANGUAGE SPOKEN AT HOME FOR THE POPULATION 25 YEARS AND OVER
* RACE - B02001_002 
    + Estimate!!Total:!!White alone RACE
* POVERY - B17001_002 
    + Estimate!!Total:!!Income in the past 12 months below poverty level: POVERTY STATUS IN THE PAST 12 MONTHS BY SEX BY AGE
* INCOME - B08121_001 
    + Estimate!!Median earnings in the past 12 months --!!Total:MEDIAN EARNINGS IN THE PAST 12 MONTHS (IN 2021 
* Employment - B23025_002 
    + Estimate!!Total:!!In labor force: EMPLOYMENT STATUS FOR THE POPULATION 16 YEARS AND OVER
```{r}
#| label: ACS
#| echo: false
#| include: false


#Population - B01003_001 Estimate!!Total TOTAL POPULATION
#HISPANIC - B03001_003 Estimate!!Total:!!Hispanic or Latino: HISPANIC OR LATINO ORIGIN BY SPECIFIC ORIGIN tract
#EDUCATION - Sum of the following
#maybe use this as the total
# B16010_001 Estimate!!Total:EDUCATIONAL ATTAINMENT AND EMPLOYMENT STATUS BY LANGUAGE SPOKEN AT HOME FOR THE POPULATION 25 YEARS AND OVER
# EDUCATIONAL ATTAINMENT AND EMPLOYMENT STATUS BY LANGUAGE SPOKEN AT HOME FOR THE POPULATION 25 YEARS AND OVER
# B16010_015 Estimate!!Total:!!High school graduate (includes equivalency): EDUCATIONAL ATTAINMENT AND EMPLOYMENT STATUS BY LANGUAGE SPOKEN AT HOME FOR THE POPULATION 25 YEARS AND OVER
# B16010_028 Estimate!!Total:!!Some college or associate's degree:EDUCATIONAL ATTAINMENT AND EMPLOYMENT STATUS BY LANGUAGE SPOKEN AT HOME FOR THE POPULATION 25 YEARS AND OVER
# B16010_041 Estimate!!Total:!!Bachelor's degree or higher: EDUCATIONAL ATTAINMENT AND EMPLOYMENT STATUS BY LANGUAGE SPOKEN AT HOME FOR THE POPULATION 25 YEARS AND OVER

#RACE - B02001_002 Estimate!!Total:!!White alone RACE
#POVERY - B17001_002 Estimate!!Total:!!Income in the past 12 months below poverty level: POVERTY STATUS IN THE PAST 12 MONTHS BY SEX BY AGE
# B17001_001 Estimate!!Total:POVERTY STATUS IN THE PAST 12 MONTHS BY SEX BY AGE
#INCOME - B08121_001 Estimate!!Median earnings in the past 12 months --!!Total:MEDIAN EARNINGS IN THE PAST 12 MONTHS (IN 2021 
#Employment - 
#probably should not use, total in labor force, includes armed forces and those that are unemployed
# B23025_002 Estimate!!Total:!!In labor force: EMPLOYMENT STATUS FOR THE POPULATION 16 YEARS AND OVER
# B23025_003 Estimate!!Total:!!In labor force:!!Civilian labor force: EMPLOYMENT STATUS FOR THE POPULATION 16 YEARS AND OVER
# B23025_004 Estimate!!Total:!!In labor force:!!Civilian labor force:!!Employed EMPLOYMENT STATUS FOR THE POPULATION 16 YEARS AND OVER
# B23025_001 Estimate!!Total:EMPLOYMENT STATUS FOR THE POPULATION 16 YEARS AND OVER

# x = load_variables(2021, "acs5", cache = TRUE)
# x %>% View()

zipcode.df <- get_acs(geography = "zcta",
                      variables = c("B01003_001",
                                    "B03001_003",
                                    "B16010_001",
                                    "B16010_015",
                                    "B16010_028",
                                    "B16010_042",
                                    "B02001_002",
                                    "B17001_001",
                                    "B17001_002",
                                    "B08121_001",
                                    "B23025_003",
                                    'B23025_004'),
                      year = 2021,
                      survey = "acs5",
                      zcta = county_zip_list) 

zipcode_wide.df = zipcode.df %>%
  select(GEOID, variable, estimate) %>%
  pivot_wider(names_from = variable,
              values_from = estimate) %>%
  rename(zc_population = B01003_001,
         zc_hispanic = B03001_003, 
         zc_education_total = B16010_001,
         zc_education_high_school = B16010_015, 
         zc_education_some_college = B16010_028, 
         zc_education_bach_plus = B16010_042, 
         zc_white = B02001_002,
         zc_poverty = B17001_002,
         zc_poverty_total = B17001_001,
         zc_income = B08121_001,
         zc_employment = B23025_004,
         zc_employment_total = B23025_003) %>%
  mutate(zc_education = zc_education_some_college + zc_education_bach_plus+zc_education_high_school,
         zc_income = zc_income/1000) %>%
   mutate(predictor_zc_per_poverty = zc_poverty / zc_population) %>%
  mutate(predictor_zc_per_hispanic = zc_hispanic / zc_population) %>%
  mutate(predictor_zc_per_education = zc_education / zc_education_total) %>%
  mutate(predictor_zc_per_race = zc_white / zc_population) %>%
  mutate(predictor_zc_per_employment = zc_employment / zc_employment_total)

#create binary versions (high/low) of zip code level predictors based on median values
# zipcode_wide.df <- zipcode_wide.df %>%
#   mutate(across(.cols = c(predictor_zc_per_poverty,
#                           predictor_zc_per_hispanic,
#                           predictor_zc_per_education,
#                           predictor_zc_per_race,
#                           predictor_zc_per_employment), .fns = function(x){
#                             med.x = median(x, na.rm = TRUE)
#                             binary = as.numeric(x > med.x)
#                             return(binary)
#                           }))
#                 # , .names = "binary_{.col}"))

county.df = left_join(county.df, zipcode_wide.df,
                      by = c( "rsd_zip_cd" = "GEOID")) 

county.df = left_join(county.df, zipcode_wide.df,
                      by = c( "cur_zip_cd" = "GEOID"), suffix = c("_rsd", "_cur"))
                



```



# Research question

Research Question 3: Identify predictors (including viral clustering) that are associated with the following outcomes

* Stage 3-4 new diagnoses. 
* Viral suppression in 2022
* In-care in 2022

## Outcome generation

### Late stage diagnosis

We dichotomize the HIV stage of an individual.

* HIV ONLY (HIV, STAGE 1 OR 2) 

VS

* HIV AND AIDS SIMULTANEOUSLY (HIV, STAGE 3)
* HIV AND LATER AIDS (HIV, STAGE 3)


### Viral suppression

We dichotomize the viral suppresison status of an individual in the calendar year 2022. As recommended by county public health, we included viral loads from the previous quarter of 2021 in the calculation of viral suppression status for 2022 (Oct, Nov, Dec 2021). If no viral load was available, the individual was considered not virally suppressed.

* Virally suppressed

VS

* Not virally suppressed



### In-care

We dichotomize the in-care status of an individual in the calendar year 2022. As recommended by county public health, we included viral loads from the previous quarter of 2021 in the calculation of viral suppression status for 2022 (Oct, Nov, Dec 2021). We used the following labs as evidence of being in-care

* CD4
* Viral load
* Genotype


* In-care

VS

* Not in-care


```{r}
#| label: outcome-generation
#| echo: false
#| include: false

################
## Generate outcomes
################

county.df = county.df %>%
  mutate(outcome_diagnosis = case_when(
    `Disease Category` == "HIV ONLY (HIV, STAGE 1 OR 2)" ~ 0,
    `Disease Category` == "HIV AND LATER AIDS (HIV, STAGE 3)" ~ 0,
    `Disease Category` == "HIV AND AIDS SIMULTANEOUSLY (HIV, STAGE 3)" ~ 1,
    TRUE ~ NA_real_),
    outcome_diagnosis_factor = factor(outcome_diagnosis,levels = c(0,1), labels = c('Early Stage','Late Stage')))

county.df = county.df %>%
  mutate(outcome_care = In_care_2022_COC,
         outcome_care_factor = factor(outcome_care,levels = c(0,1), labels = c('Not in-care','In-care'))) 

county.df = county.df %>%
  mutate(outcome_TEMP = VL_supp_2022_COC) %>%
  mutate(outcome_vl_supp = case_when(
    outcome_TEMP == 0 ~ 0,
    outcome_TEMP == 1 ~ 1,
    TRUE ~ 0),
    outcome_vl_supp_factor = factor(outcome_vl_supp,levels = c(0,1), labels = c('Unsuppressed','Suppressed')))

#filter out one person with missing outcome for late stage dx
county.df <- county.df %>%
  filter(!is.na(outcome_diagnosis) & !is.na(outcome_care) & !is.na(outcome_vl_supp))


```

# Predictors

Individual-level variables to consider: 


*   predictor_diagnosis_year
*   Age.diagnosis
*   predictor_hispanic
*   predictor_idu
*   predictor_msm
<!-- *   predictor_education -->
*   predictor_birth_sex
*   clustered_2022

zip code-level variables to consider: 


*   Median income
*   Percent in poverty
*   Percent clustered in 2022
*   predictor_zc_per_poverty
*   predictor_zc_per_hispanic
*   predictor_zc_per_education
*   predictor_zc_per_race
*   predictor_zc_per_employment

```{r}
#| label: predictors
#| echo: false
#| include: false

################
## Predictors
################

county.df = county.df %>%
  mutate(predictor_education = case_when(
    Education == "<= 8TH GRADE" ~ 0,
    Education == "SOME SCHOOL, LEVEL UNKNOWN" ~ 0,
    Education == "SOME HIGH SCHOOL" ~ 0,
    Education == "HIGH SCHOOL GRAD" ~ 0,
    Education == "SOME COLLEGE" ~ 1,
    Education == "COLLEGE DEGREE" ~ 1,
    Education == "POST-GRADUATE WORK" ~ 1,
    TRUE ~ NA_real_))

county.df = county.df %>%
  mutate(predictor_hispanic = case_when(
    Race == "WHITE" ~ 0,
    Race == "OTHER" ~ 0,
    Race == "BLACK" ~ 0,
    Race == "HISPANIC, ANY RACE" ~ 1,
    TRUE ~ NA_real_))

county.df = county.df %>%
  mutate(predictor_msm = case_when(
    `Exposure Category` == "MSM" ~ 1,
    `Exposure Category` == "IDU" ~ 0,
    `Exposure Category` == "NO REPORTED RISK" ~ 0,
    `Exposure Category` == "MSM & IDU" ~ 1,
    `Exposure Category` == "HETEROSEXUAL CONTACT" ~ 0,
    `Exposure Category` == "OTHER" ~ 0,
    `Exposure Category` == "PERINATAL EXPOSURE" ~ 0,
    TRUE ~ NA_real_))

county.df = county.df %>%
  mutate(predictor_idu = case_when(
    `Exposure Category` == "MSM" ~ 0,
    `Exposure Category` == "IDU" ~ 1,
    `Exposure Category` == "NO REPORTED RISK" ~ 0,
    `Exposure Category` == "MSM & IDU" ~ 1,
    `Exposure Category` == "HETEROSEXUAL CONTACT" ~ 0,
    `Exposure Category` == "OTHER" ~ 0,
    `Exposure Category` == "PERINATAL EXPOSURE" ~ 0,
    TRUE ~ NA_real_))

county.df = county.df %>%
  mutate(predictor_diagnosis_year = `Diagnosis year`)

county.df = county.df %>%
  mutate(predictor_birth_sex = case_when(
    `Birth Sex` == "FEMALE" ~ 0,
    `Birth Sex` == "MALE" ~ 1,
    TRUE ~ NA_real_))


    
reg_variable_ind.vec = c(
  "predictor_diagnosis_year",
  "Age.diagnosis",
  "predictor_hispanic",
  "predictor_idu",
  "predictor_msm",
  #removing education since there are over 1,000 people with a missing value for education
  # "predictor_education",
  "predictor_birth_sex",
  "clustered_2022")

reg_variable_zip.vec = c(
  "percent_clustered_2022",
  "zc_income",
  "predictor_zc_per_poverty",
  "predictor_zc_per_hispanic",
  "predictor_zc_per_education",
  "predictor_zc_per_race",
  "predictor_zc_per_employment"
)

reg_variable_rsd_zip.vec = paste(reg_variable_zip.vec, "rsd", sep="_")
reg_variable_cur_zip.vec = paste(reg_variable_zip.vec, "cur", sep="_")

# reg_variable_rsd_zip.vec.cont <- reg_variable_rsd_zip.vec
# reg_variable_cur_zip.vec.cont <- reg_variable_cur_zip.vec

# #create special list of cars with continuous versions to check NA and rmeove people with missing values
# reg_variable.vec_rsd_checkna = c(reg_variable_ind.vec, reg_variable_rsd_zip.vec.cont)
# reg_variable.vec_cur_checkna = c(reg_variable_ind.vec, reg_variable_cur_zip.vec.cont)
# 
reg_variable.vec_rsd = c(reg_variable_ind.vec, reg_variable_rsd_zip.vec)
reg_variable.vec_cur = c(reg_variable_ind.vec, reg_variable_cur_zip.vec)

################
## Add testing data
################


if (testing_site_bool) {
  
  # county_testing_site.df = read_excel(paste(directory_loc, county_test_file_subloc, sep="")) %>%
  #   rename(zip_code = `Site zip code`) %>%
  #   mutate(across(where(is.character), toupper)) %>%
  #   mutate(zip_code = pad.zip(zip_code),
  #          testing_site_bin = 1) %>%
  #   select(zip_code, testing_site_bin) 
  
  county_testing_site.df = read_excel(paste(directory_loc, county_test_file_subloc, sep="")) %>%
    rename(zip_code = `Site zip code`) %>%
    mutate(across(where(is.character), toupper)) %>%
    mutate(zip_code = pad.zip(zip_code)) %>%
    group_by(zip_code) %>%
    summarize(n.testing.sites.zip = n(),
     testing_site_bin = 1) %>%
    select(zip_code, testing_site_bin,n.testing.sites.zip)
  
  county.df = county.df %>% 
    left_join(county_testing_site.df,
              by = join_by(rsd_zip_cd == zip_code),
              relationship = "many-to-one") %>%
      left_join(county_testing_site.df,
              by = join_by(cur_zip_cd == zip_code),
              relationship = "many-to-one",suffix = c('_rsd','_cur')) %>%
    mutate(across(.cols = c(testing_site_bin_rsd,testing_site_bin_cur),.fn = ~ifelse(is.na(.x), 0, .x)))
  
  reg_variable.vec_rsd <- c(reg_variable.vec_rsd, "testing_site_bin_rsd")
  reg_variable.vec_cur <- c(reg_variable.vec_cur, "testing_site_bin_cur")
  
}

#save a copy of the data with missing zip code level characterics so I can calculate number missing and include in methods
county.df.includeNA <- county.df

#filter out people with missing zip code level characterics due to values missing from ACS
county.df = county.df %>%
  filter(!if_any(.cols = all_of(c(reg_variable.vec_rsd,reg_variable.vec_cur)),~ is.na(.)))

# county.df.cur = county.df %>%
#   filter(!if_any(.cols = all_of(reg_variable.vec_cur),~ is.na(.)))

#convert zip code level variables to high/low using median after all people with misisng data have been removed
county.df <- county.df %>%
  mutate(across(.cols = c(matches('predictor_zc|percent_clustered|zc_income')), .fns = function(x){
                            med.x = median(x, na.rm = TRUE)
                            binary = as.numeric(x > med.x)
                            # binary = as.numeric(x >= med.x)
                            # return(binary)
                            return(data.frame(binary = binary,med.x = med.x))
                          },.unpack = T))


#paste binary onto zip code variable names to use high/low versions just created above
reg_variable_cur_zip.vec <- paste(reg_variable_cur_zip.vec, "binary", sep="_")
reg_variable_rsd_zip.vec <- paste(reg_variable_rsd_zip.vec, "binary", sep="_")



reg_variable.vec_rsd <- gsub(x = reg_variable.vec_rsd, pattern = '^((?!test).*)(_rsd)$', replacement = '\\1\\2_binary',fixed = F,perl = T)
reg_variable.vec_cur <- gsub(x = reg_variable.vec_cur, pattern = '^((?!test).*)(_cur)$', replacement = '\\1\\2_binary',fixed = F,perl = T)

county.df.rsd_zip_cd.single = county.df %>%
  distinct(rsd_zip_cd,.keep_all = TRUE) %>%
  select(rsd_zip_cd, 
         all_of(reg_variable_rsd_zip.vec))

county.df.cur_zip_cd.single = county.df %>%
  distinct(cur_zip_cd,.keep_all = TRUE) %>%
  select(cur_zip_cd, 
         all_of(reg_variable_cur_zip.vec))

```

Number of individuals after filtering to diagnosed in county: `r county.df %>% nrow()`

# Descriptive statistics

The number of individuals in Stages 1 or 2 (outcome = 0) vs. number of individuals in Stage 3

```{r}
#| label: descriptive-statistics
#| echo: false
#| include: false

################
## Table 1
################

county.tbl1 = county.df %>% 
  select(Age.diagnosis, predictor_diagnosis_year,Race.factor, `Birth Sex.factor`, `Exposure Category`,
         clustered_2022,Sequence_ever,all_of(reg_variable_cur_zip.vec)) %>% # keep only columns of interest
  tbl_summary(     
    # by = outcome,                                               # stratify entire table by outcome
    statistic = list(all_continuous() ~ "{mean} ({p0}, {p100})",        # stats and format for continuous columns
                     all_categorical() ~ "{n} ({p}%)"),   # stats and format for categorical columns
                     # all_categorical() ~ "{n} / {N} ({p}%)"),   # stats and format for categorical columns
    digits = all_continuous() ~ 1,                              # rounding for continuous columns
    type   = all_categorical() ~ "categorical",                 # force all categorical levels to display
    label  = list(                                              # display labels for column names
       Age.diagnosis ~ "Age at diagnosis",
        predictor_diagnosis_year ~ "Year of diagnosis",
      Race.factor   ~ "Race",                           
      `Birth Sex.factor` ~ "Sex",
      `Exposure Category` ~ "Transmission Risk",
      # Education ~ "Education",
      Sequence_ever ~ 'Sequence available',
      clustered_2022 ~ "Genetic cluster",
      percent_clustered_2022_cur_binary ~ "Percentage in cluster",
        zc_income_cur_binary ~ "Median income",
       predictor_zc_per_poverty_cur_binary ~ "Percentage below poverty level",
      predictor_zc_per_hispanic_cur_binary ~ "Percentage Hispanic",
      predictor_zc_per_education_cur_binary ~ 'Percentage high school education',
      predictor_zc_per_race_cur_binary ~ 'Percentage White',
      predictor_zc_per_employment_cur_binary ~ 'Percentage employed'),
    missing_text = "Missing"                                    # how missing values should display
  )

write_csv(county.tbl1 %>% as.data.frame(), paste0(result_file_loc, "county_tbl1.csv", sep = ""))

outcomes.factor.list <- list("diagnosis" = list(outcome = 'outcome_diagnosis_factor',
                              random = 'rsd_zip_cd',
                              data = county.df,
                              predictors.zip = reg_variable_rsd_zip.vec,
                              suffix.zip = '_rsd_binary'),
                         'care' = list(outcome = 'outcome_care_factor',
                              random = 'cur_zip_cd',
                              data = county.df,
                              predictors.zip = reg_variable_cur_zip.vec,
                              suffix.zip = '_cur_binary'),
                         'supp' = list(outcome = 'outcome_vl_supp_factor',
                                       random = 'cur_zip_cd',
                                       data = county.df,
                                       predictors.zip = reg_variable_cur_zip.vec,
                              suffix.zip = '_cur_binary'))
# 
outcomes.factor <- c('outcome_diagnosis_factor', 'outcome_care_factor', 'outcome_vl_supp_factor')
# outcomes.factor.list <- as.list(outcomes.factor)
names(outcomes.factor.list) <- outcomes.factor

outcome <- outcomes.factor.list[[1]]

list.tab1s <- lapply(outcomes.factor.list,FUN = function(outcome){
  county.tbl1.split = outcome$data %>% 
    select(Age.diagnosis, predictor_diagnosis_year,Race.factor, `Birth Sex.factor`, `Exposure Category`,
         clustered_2022,Sequence_ever,all_of(outcome$predictors.zip),
           !!outcome$outcome) %>% # keep only columns of interest
    # select(Age.diagnosis, predictor_diagnosis_year,Race.factor, `Birth Sex.factor`, `Exposure Category`,
    #        clustered_2022,Sequence_ever,all_of(reg_variable_cur_zip.vec),
           # !!outcome$outcome) %>% # keep only columns of interest
    rename_with(~ str_replace(., outcome$suffix.zip, ""),ends_with(outcome$suffix.zip)) %>%
    tbl_summary(     
      by = !!outcome$outcome,                                               # stratify entire table by outcome
      statistic = list(all_continuous() ~ "{mean} ({p0}, {p100})",        # stats and format for continuous columns
                       all_categorical() ~ "{n} ({p}%)"),   # stats and format for categorical columns
                       # all_categorical() ~ "{n} / {N} ({p}%)"),   # stats and format for categorical columns
      digits = all_continuous() ~ 1,                              # rounding for continuous columns
      type   = all_categorical() ~ "categorical",                 # force all categorical levels to display
      label  = list(                                              # display labels for column names
        Age.diagnosis ~ "Age at diagnosis",
        predictor_diagnosis_year ~ "Year of diagnosis",
        Race.factor   ~ "Race",                           
        `Birth Sex.factor` ~ "Sex",
        `Exposure Category` ~ "Transmission Risk",
        # Education ~ "Education",
        Sequence_ever ~ 'Sequence available',
        clustered_2022 ~ "Genetic cluster",
        percent_clustered_2022 ~ "Percentage in cluster",
        zc_income ~ "Median income",
       predictor_zc_per_poverty ~ "Percentage below poverty level",
      predictor_zc_per_hispanic ~ "Percentage Hispanic",
      predictor_zc_per_education ~ 'Percentage high school education',
      predictor_zc_per_race ~ 'Percentage White',
      predictor_zc_per_employment ~ 'Percentage employed'),
        # outcome_diagnosis  ~ "Late-stage Diagnosis",
        # outcome_care  ~ "In Care",
        # outcome_vl_supp  ~ "Viral Suppression"),
      missing_text = "Missing"                                    # how missing values should display
    ) %>%
  add_overall()
  write_csv(county.tbl1.split %>% as.data.frame(), paste0(result_file_loc, "county_tbl1_",outcome$outcome,".csv", sep = ""))
  return(list(html.table= county.tbl1.split,
         df.table = county.tbl1.split %>% as.data.frame() %>% select(-contains("Overall"))))
})

county.tbl1.full <- bind_cols(bind_cols(list.tab1s$outcome_vl_supp_factor$df.table,
                                        list.tab1s$outcome_care_factor$df.table,
                                        list.tab1s$outcome_diagnosis_factor$df.table),
                              county.tbl1 %>% as.data.frame()) %>%
  select(-c("**Characteristic**...4" ,"**Characteristic**...7","**Characteristic**...10" ))

write_csv(county.tbl1.full %>% as.data.frame(), paste0(result_file_loc, "county_tbl1_full.csv", sep = ""))

#can create pseudo dataset and use table strata to create one table as well
# tbl_strata_ex1 <- county.df %>% 
#   select(Race.factor, `Birth Sex.factor`, `Exposure Category`, Education,
#          clustered_2022,Sequence_ever,
#          outcome_diagnosis_factor,outcome_care_factor) %>% # keep only columns of interest
#   tbl_strata(
#     strata = c(outcome_diagnosis_factor,outcome_care_factor),
#     .tbl_fun =
#       ~ .x %>%
#       tbl_summary(     
#         # by = !!outcome,                                               # stratify entire table by outcome
#         statistic = list(all_continuous() ~ "{mean} ({sd})",        # stats and format for continuous columns
#                          all_categorical() ~ "{n} / {N} ({p}%)"),   # stats and format for categorical columns
#         digits = all_continuous() ~ 1,                              # rounding for continuous columns
#         type   = all_categorical() ~ "categorical",                 # force all categorical levels to display
#         label  = list(                                              # display labels for column names
#           Race.factor   ~ "Race",                           
#           `Birth Sex.factor` ~ "Sex",
#           `Exposure Category` ~ "Transmission Risk",
#           Education ~ "Education",
#           Sequence_ever ~ 'Sequence available',
#           clustered_2022 ~ "Genetic cluster"),
#         missing_text = "Missing") %>%
#       add_n(),
#     .header = "**{strata}**, N = {n}"
#     # .header = "**{strata}**"
#   )



################
## zip code descriptive tables
################

county.rsd_zip.tbl1 = county.df %>% 
  select(all_of(reg_variable_rsd_zip.vec)) %>% # keep only columns of interest
  tbl_summary(     
    # by = outcome,                                               # stratify entire table by outcome
    statistic = list(#all_continuous() ~ "{mean} ({sd})",        # stats and format for continuous columns
                     all_continuous() ~ "{mean} ({p0}, {p100}))",        # stats and format for continuous columns
                     all_categorical() ~ "{n} / {N} ({p}%)"),   # stats and format for categorical columns
    digits = all_continuous() ~ 1,                              # rounding for continuous columns
    type   = all_categorical() ~ "categorical",                 # force all categorical levels to display
    label  = list(                                              # display labels for column names
      percent_clustered_2022_rsd_binary   ~ "Percent clustered (2022)",                           
      zc_income_rsd_binary ~ "Median income",
      predictor_zc_per_poverty_rsd_binary ~ "Percentage below poverty level",
      predictor_zc_per_hispanic_rsd_binary ~ "Percentage Hispanic",
      predictor_zc_per_education_rsd_binary ~ 'Percentage high school education or higher',
      predictor_zc_per_race_rsd_binary ~ 'Percentage White',
      predictor_zc_per_employment_rsd_binary ~ 'Percentage employed'),
    missing_text = "Missing"                                    # how missing values should display
  )
write_csv(county.rsd_zip.tbl1 %>% as.data.frame(), paste0(result_file_loc, "county_tbl1_rsd_zip.csv", sep = ""))

county.cur_zip.tbl1 = county.df %>% 
  select(all_of(reg_variable_cur_zip.vec)) %>% # keep only columns of interest
  tbl_summary(     
    # by = outcome,                                               # stratify entire table by outcome
    statistic = list(#all_continuous() ~ "{mean} ({sd})",        # stats and format for continuous columns
                     all_continuous() ~ "{mean} ({p0}, {p100}))",        # stats and format for continuous columns
                     all_categorical() ~ "{n} / {N} ({p}%)"),   # stats and format for categorical columns
    digits = all_continuous() ~ 1,                              # rounding for continuous columns
    type   = all_categorical() ~ "categorical",                 # force all categorical levels to display
    label  = list(                                              # display labels for column names
      percent_clustered_2022_cur_binary   ~ "Percent clustered (2022)",
      zc_income_cur_binary ~ "Median income",
      predictor_zc_per_poverty_cur_binary ~ "Percentage below poverty level",
      predictor_zc_per_hispanic_cur_binary ~ "Percentage Hispanic",
      predictor_zc_per_education_cur_binary ~ 'Percentage high school education or higher',
      predictor_zc_per_race_cur_binary ~ 'Percentage White',
      predictor_zc_per_employment_cur_binary ~ 'Percentage employed'),
    missing_text = "Missing"                                    # how missing values should display
  )
write_csv(county.cur_zip.tbl1 %>% as.data.frame(), paste0(result_file_loc, "county_tbl1_cur_zip.csv", sep = ""))

# county.df.includeNA.cc <- county.df.includeNA %>%
#   select("zc_income", "predictor_zc_per_poverty", "predictor_zc_per_hispanic"  ,"predictor_zc_per_education",  "predictor_zc_per_race", "predictor_zc_per_employment") %>% 
#   filter(!complete.cases(.))
#   na.omit()
# county.zip.tbl1.includeNA = county.df.includeNA %>% 
#   select(reg_variable_zip.vec) %>% # keep only columns of interest
#   tbl_summary(     
#     # by = outcome,                                               # stratify entire table by outcome
#     statistic = list(#all_continuous() ~ "{mean} ({sd})",        # stats and format for continuous columns
#                      all_continuous() ~ "{mean} ({p0}, {p100}))",        # stats and format for continuous columns
#                      all_categorical() ~ "{n} / {N} ({p}%)"),   # stats and format for categorical columns
#     digits = all_continuous() ~ 1,                              # rounding for continuous columns
#     type   = all_categorical() ~ "categorical",                 # force all categorical levels to display
#     label  = list(                                              # display labels for column names
#       percent_clustered_2022   ~ "Percent clustered (2022)",                           
#       zc_income ~ "Median income",
#       predictor_zc_per_poverty ~ "Percentage below poverty level",
#       predictor_zc_per_hispanic ~ "Percentage Hispanic",
#       predictor_zc_per_education ~ 'Percentage high school education or higher',
#       predictor_zc_per_race ~ 'Percentage White',
#       predictor_zc_per_employment ~ 'Percentage employed'),
#     missing_text = "Missing"                                    # how missing values should display
#   )

```

```{r}
#| echo: false
#| output: true
#| tbl-cap: Overall population included in analysis. Zip code level variables shown are from the zip code at diagnosis
county.tbl1
```

## Late stage diagnosis

```{r}
#| echo: false
#| output: true
#| tbl-cap: Late stage diagnosis.Zip code level variables shown are from the zip code at diagnosis
list.tab1s$outcome_diagnosis_factor$html.table
```

## Virally suppressed

```{r}
#| echo: false
#| output: true
#| tbl-cap: Virally suppressed. Zip code level variables shown are from the zip code of current residence
list.tab1s$outcome_vl_supp_factor$html.table
```

## In-care

```{r}
#| echo: false
#| output: true
#| tbl-cap: In-care. Zip code level variables shown are from the zip code of current residence
list.tab1s$outcome_care_factor$html.table
```

## Zip-code level variables

### Zip code at diagnosis

```{r}
#| echo: false
#| output: true
#| tbl-cap: rsdzip-predictors
county.rsd_zip.tbl1
```

### Zip code of current residence


```{r}
#| echo: false
#| output: true
#| tbl-cap: curzip-predictors
county.cur_zip.tbl1
```



# Regression

```{r}
#| label: regression
#| echo: false
#| include: false

########
########
# 
# county.df = county.df %>%
#   mutate(med_percent_clustered_2022_rsd = median(percent_clustered_2022_rsd)) %>% 
#   mutate(percent_clustered_2022_rsd = as.numeric(percent_clustered_2022_rsd > med_percent_clustered_2022_rsd)) %>%
#   select(-med_percent_clustered_2022_rsd)
# 
# county.df = county.df %>%
#   mutate(med_predictor_zc_per_poverty_rsd = median(predictor_zc_per_poverty_rsd)) %>% 
#   mutate(predictor_zc_per_poverty_rsd = as.numeric(predictor_zc_per_poverty_rsd > med_predictor_zc_per_poverty_rsd)) %>%
#   select(-med_predictor_zc_per_poverty_rsd)
# 
# county.df = county.df %>%
#   mutate(med_zc_income_rsd = median(zc_income_rsd)) %>% 
#   mutate(zc_income_rsd = as.numeric(zc_income_rsd > med_zc_income_rsd)) %>%
#   select(-med_zc_income_rsd)
# 
# county.df = county.df %>%
#   mutate(med_predictor_zc_per_hispanic_rsd = median(predictor_zc_per_hispanic_rsd)) %>% 
#   mutate(predictor_zc_per_hispanic_rsd = as.numeric(predictor_zc_per_hispanic_rsd > med_predictor_zc_per_hispanic_rsd)) %>%
#   select(-med_predictor_zc_per_hispanic_rsd)
# 
# county.df = county.df %>%
#   mutate(med_predictor_zc_per_education_rsd = median(predictor_zc_per_education_rsd)) %>% 
#   mutate(predictor_zc_per_education_rsd = as.numeric(predictor_zc_per_education_rsd > med_predictor_zc_per_education_rsd)) %>%
#   select(-med_predictor_zc_per_education_rsd)
# 
# county.df = county.df %>%
#   mutate(med_predictor_zc_per_race_rsd = median(predictor_zc_per_race_rsd)) %>% 
#   mutate(predictor_zc_per_race_rsd = as.numeric(predictor_zc_per_race_rsd > med_predictor_zc_per_race_rsd)) %>%
#   select(-med_predictor_zc_per_race_rsd)
# 
# county.df = county.df %>%
#   mutate(med_predictor_zc_per_employment_rsd = median(predictor_zc_per_employment_rsd)) %>% 
#   mutate(predictor_zc_per_employment_rsd = as.numeric(predictor_zc_per_employment_rsd > med_predictor_zc_per_employment_rsd)) %>%
#   select(-med_predictor_zc_per_employment_rsd)
# 
# ###
# 
# county.df = county.df %>%
#   mutate(med_percent_clustered_2022_cur = median(percent_clustered_2022_cur)) %>% 
#   mutate(percent_clustered_2022_cur = as.numeric(percent_clustered_2022_cur > med_percent_clustered_2022_cur)) %>%
#   select(-med_percent_clustered_2022_cur)
# 
# county.df = county.df %>%
#   mutate(med_predictor_zc_per_poverty_cur = median(predictor_zc_per_poverty_cur)) %>% 
#   mutate(predictor_zc_per_poverty_cur = as.numeric(predictor_zc_per_poverty_cur > med_predictor_zc_per_poverty_cur)) %>%
#   select(-med_predictor_zc_per_poverty_cur)
# 
# county.df = county.df %>%
#   mutate(med_zc_income_cur = median(zc_income_cur)) %>% 
#   mutate(zc_income_cur = as.numeric(zc_income_cur > med_zc_income_cur)) %>%
#   select(-med_zc_income_cur)
# 
# county.df = county.df %>%
#   mutate(med_predictor_zc_per_hispanic_cur = median(predictor_zc_per_hispanic_cur)) %>% 
#   mutate(predictor_zc_per_hispanic_cur = as.numeric(predictor_zc_per_hispanic_cur > med_predictor_zc_per_hispanic_cur)) %>%
#   select(-med_predictor_zc_per_hispanic_cur)
# 
# county.df = county.df %>%
#   mutate(med_predictor_zc_per_education_cur = median(predictor_zc_per_education_cur)) %>% 
#   mutate(predictor_zc_per_education_cur = as.numeric(predictor_zc_per_education_cur > med_predictor_zc_per_education_cur)) %>%
#   select(-med_predictor_zc_per_education_cur)
# 
# county.df = county.df %>%
#   mutate(med_predictor_zc_per_race_cur = median(predictor_zc_per_race_cur)) %>% 
#   mutate(predictor_zc_per_race_cur = as.numeric(predictor_zc_per_race_cur > med_predictor_zc_per_race_cur)) %>%
#   select(-med_predictor_zc_per_race_cur)
# 
# county.df = county.df %>%
#   mutate(med_predictor_zc_per_employment_cur = median(predictor_zc_per_employment_cur)) %>% 
#   mutate(predictor_zc_per_employment_cur = as.numeric(predictor_zc_per_employment_cur > med_predictor_zc_per_employment_cur)) %>%
#   select(-med_predictor_zc_per_employment_cur)



########
########

# reg_variable.vec_rsd = reg_variable.vec_rsd[-which(reg_variable.vec_rsd =="zc_income_rsd_binary")]
# reg_variable.vec_cur = reg_variable.vec_cur[-which(reg_variable.vec_cur =="zc_income_cur_binary")]

################
## Regression
################
# 
outcome_var_list = c("outcome_diagnosis", "outcome_care", "outcome_vl_supp")
random_effect_var_list = c("rsd_zip_cd", "cur_zip_cd", "cur_zip_cd")

outcome_var_list <- list("diagnosis" = list(outcome = 'outcome_diagnosis',
                              random = 'rsd_zip_cd',
                              data = county.df,
                              predictors = reg_variable.vec_rsd),
                         'care' = list(outcome = 'outcome_care',
                              random = 'cur_zip_cd',
                              data = county.df,
                              predictors = reg_variable.vec_cur),
                         'supp' = list(outcome = 'outcome_vl_supp',
                                       random = 'cur_zip_cd',
                                       data = county.df,
                                       predictors = reg_variable.vec_cur))

var_list <- outcome_var_list[[1]]

regression.output.list <- lapply(X = outcome_var_list,FUN = function(var_list){

  outcome_var_i = var_list$outcome
  print(outcome_var_i)

  desc_res.df = var_list$data %>%
    select(all_of(var_list$predictors), !! outcome_var_i) %>%
    tbl_summary(by = !! outcome_var_i,
                digits = all_continuous() ~ 1,  # display 1 decimal for continuous variables
                 statistic = list(#all_continuous() ~ "{mean} ({sd})",        # stats and format for continuous columns
                     all_continuous() ~ "{mean} ({p0}, {p100}))",        # stats and format for continuous columns
                     all_categorical() ~ "{n} / {N} ({p}%)")) %>% as_tibble() %>%
    rename(variable = `**Characteristic**`)

  print('univariate')
  uni_reg_res.df = univariate_reg(county.df = var_list$data,
                                  predictors = var_list$predictors,
                                  outcome_var = var_list$outcome,
                                  random_effect_var = var_list$random)
  print('lasso tune')
  lambda_tune = Lasso_tune_lambda(county.df = var_list$data,
                                  predictors = var_list$predictors,
                                  outcome_var = var_list$outcome,
                                  random_effect_var = var_list$random)
  print('lasso reg')
  multi_reg_res.df = Lasso_reg(county.df = var_list$data,
                               predictors = var_list$predictors,
                               outcome_var = var_list$outcome,
                               random_effect_var = var_list$random,
                               lambda_tune = lambda_tune)

  reg_res.df = full_join(uni_reg_res.df$summary.table, multi_reg_res.df$summary.table, by = "variable")

  outcome_res.df = left_join(desc_res.df, reg_res.df, by = "variable")
  outcome_res.df.round <- outcome_res.df %>%
    mutate(across(.cols = c(estimate_uni, estimate_multi),.fns = function(x){round(x,digits = 2)})) %>%
    mutate(across(.cols = c(p_val_uni,p_val_multi ),.fns = function(x){pvalFormat(x,empty.cell.value = '-')}))
  
  outcome_res.df.round.paper <- outcome_res.df.round %>%
    mutate(estimate_uni_se = paste0(estimate_uni, " (", ifelse(se_uni < 1e-3,
                                                               formatC(se_uni, format = "E", digits = 2), 
                                                               round(se_uni,3)), ")"),
          estimate_multi_se = paste0(estimate_multi, " (", ifelse(se_multi < 1e-3,
                                                               formatC(se_multi, format = "E", digits = 2), 
                                                               round(se_multi,3)), ")")) %>%
    relocate(estimate_uni_se, .before = estimate_uni) %>%
    relocate(estimate_multi_se, .before = estimate_multi)

  write_csv(outcome_res.df, paste0(result_file_loc, outcome_var_i, "_all_tbl.csv", sep = ""))
  write_csv(outcome_res.df.round, paste0(result_file_loc, outcome_var_i, "_all_tbl_rounded.csv", sep = ""))

  return(list(outcome_var = outcome_var_i,
              univariate = uni_reg_res.df,
              lambda_tune = lambda_tune,
              multi_reg_res.df = multi_reg_res.df,
              reg_res.df = reg_res.df,
              outcome_res.df = outcome_res.df,
              outcome_res.df.round = outcome_res.df.round))
})

# save(regression.output.list,file = paste0(result_file_loc, "regression.output.list.RData", sep = ""))
# load(file = paste0(result_file_loc, "regression.output.list.RData", sep = ""))
# 
# regression.output.list$supp$reg_res.df
# regression.output.list$care$reg_res.df
# regression.output.list$diagnosis$reg_res.df

tables.paper <- lapply(regression.output.list, function(output){
  outcome_res.df.round.paper <- output$outcome_res.df.round %>%
    mutate(estimate_uni_se = paste0(estimate_uni, " (", ifelse(se_uni < 1e-3,
                                                               formatC(se_uni, format = "E", digits = 2), 
                                                               round(se_uni,3)), ")"),
          estimate_multi_se = paste0(estimate_multi, " (", ifelse(se_multi < 1e-3,
                                                               formatC(se_multi, format = "E", digits = 2), 
                                                               round(se_multi,3)), ")")) %>%
    relocate(estimate_uni_se, .before = estimate_uni) %>%
    relocate(estimate_multi_se, .before = estimate_multi) %>%
    select(-c(estimate_uni,se_uni,estimate_multi,se_multi))
  return(outcome_res.df.round.paper)
})

write_csv(tables.paper$diagnosis, paste0(result_file_loc, 'diagnosis', "_all_tbl_paper_rounded_mean_range.csv", sep = ""))
write_csv(tables.paper$supp, paste0(result_file_loc, 'supp', "_all_tbl_paper_rounded_mean_range.csv", sep = ""))
write_csv(tables.paper$care, paste0(result_file_loc, 'care', "_all_tbl_paper_rounded_mean_range.csv", sep = ""))

#combine ORs into a single table with rows as predictors and columns as outcomes.
View( regression.output.list$diagnosis$outcome_res.df.round)

poster.table.prep.fx <- function(outcome_res.df.round, zip_code_suffix = '_rsd|_rsd_binary',outcome_suffix = '_diagnosis'){
  tab.out <- outcome_res.df.round %>%
  mutate(across(.cols = starts_with("p_val"),.fns = function(x){str_extract(string = x,pattern = '\\*+')},.names = '{.col}_sig')) %>%
  select(variable,estimate_uni,estimate_multi,ends_with('sig')) %>%
  rename_with(.cols = -variable,.fn = ~ paste0(.x,outcome_suffix)) %>%
  mutate(across(.cols = variable, .fn = ~ gsub(pattern = zip_code_suffix,replacement = '',x = .x)))
  
}

dx.poster.table <- poster.table.prep.fx(outcome_res.df.round = regression.output.list$diagnosis$outcome_res.df.round,
                                         zip_code_suffix = '_rsd|_rsd_binary',
                                         outcome_suffix = '_diagnosis')
care.poster.table <- poster.table.prep.fx(outcome_res.df.round = regression.output.list$care$outcome_res.df.round,
                                         zip_code_suffix = '_cur|_cur_binary',
                                         outcome_suffix = '_care')
supp.poster.table <- poster.table.prep.fx(outcome_res.df.round = regression.output.list$supp$outcome_res.df.round,
                                         zip_code_suffix = '_cur|_cur_binary',
                                         outcome_suffix = '_supp')


poster.table <- dx.poster.table %>% 
  left_join(care.poster.table, by = "variable") %>% 
  left_join(supp.poster.table, by = "variable") %>%
  mutate(across(.cols = starts_with("estimate"), .fns = function(x){ifelse(x == 0,'-',round(exp(x),2))})) %>%
  #pivot longer so easy to paste ORs and sig stars
   pivot_longer(cols = -c(variable), names_to = c(".value", "outcome"), 
                 names_pattern = "^(.*)_([a-z]+)") %>%
  mutate(across(.cols = starts_with('p_val'),.fn = ~ ifelse(is.na(.x) ,'',.x))) %>%
  mutate(estimate_uni_sig = paste0(estimate_uni,p_val_uni_sig),
         estimate_multi_sig = paste0(estimate_multi,p_val_multi_sig)) %>%
  select(variable,estimate_uni_sig,estimate_multi_sig,outcome) %>%
  #pivot back to wide
  pivot_wider(id_cols = variable,
              names_from = outcome,
    names_glue = "{.value}_{outcome}",
    values_from = c(estimate_uni_sig, estimate_multi_sig)) %>%
  relocate(variable,ends_with('diagnosis'),ends_with('care'),ends_with('supp'))

write.csv(x = poster.table, file = paste0('./manuscript/poster/', "poster.table.csv", sep = ""), row.names = FALSE)


  
```


First set of estimates and p-values are for univariate regression. The second set is for multivariate regression.

## Viral suppression

```{r}
#| echo: false
#| output: true
#| tbl-cap: Viral suppression
regression.output.list$supp$outcome_res.df.round %>%
  gt()
```


## In-care

```{r}
#| echo: false
#| output: true
#| tbl-cap: In Care
#| tbl-colwidths: auto
regression.output.list$care$outcome_res.df.round %>%
  gt()
```

## Late stage diagnosis

```{r}
#| echo: false
#| output: true
#| tbl-cap: Late stage diagnosis
regression.output.list$diagnosis$outcome_res.df.round %>%
  gt()
```

# Geographic map

```{r}
#| label: maps
#| echo: false
#| include: false

# Map function to plot intersection of poverty and in-care, still in development



outcome <- map_outcome_list[[2]]
title_a = outcome$outcome_var_list_nice
title_a = 'High vs low poverty'
county.df = county.df
county_zip_list = county_zip_list
outcome_var = outcome$outcome_var_list
geo_zip_var = outcome$random_effect_var_list
zip_var = 'predictor_zc_per_poverty_cur_binary'
min.n.obs = 10
legend.name = 'poverty'

geo_map.zipvar <- function(county.df,
                    county_zip_list,
                    zip_var,
                    geo_zip_var,
                    title_a,
                    legend.name = 'poverty',
                    min.n.obs = 10) {

  zcta_data <- get_acs(
    geography = "zcta",
    variables = "B19013_001",
    zcta = county_zip_list,
    geometry = TRUE
  )

  #for high/low census variables
  county_summary.df = county.df %>%
      group_by(!! rlang::sym(geo_zip_var)) %>%
    mutate(n_tot = n()) %>%
    ungroup() %>%
    filter(n_tot > min.n.obs) %>%
    distinct(!! rlang::sym(geo_zip_var),.keep_all = T) %>%
   select(!! rlang::sym(geo_zip_var),!! rlang::sym(zip_var),n_tot) %>%
    mutate(rsd_zip_cd_cat = as.character(!! rlang::sym(geo_zip_var)),
           predictor.zip = as.character(!! rlang::sym(zip_var)))

  zcta_level.df = left_join(zcta_data,
                            county_summary.df,
                            by = c("GEOID" = "rsd_zip_cd_cat"))

  zcta_level.df = left_join(zcta_level.df,
                            zip_code_db %>% select(zipcode, lat, lng),
                            by = c("GEOID" = "zipcode")) %>%
    filter(!is.na(lat))



  
  
  pal_fun <- colorFactor("YlOrRd", NULL) #, na.color = "#808080"
  
  p_popup <- paste0("<strong>", title_a, ": </strong>", zcta_level.df$predictor.zip, "<br>",
                    "<strong>Zip code: </strong>", zcta_level.df$GEOID)
  
  zcta_level.df = zcta_level.df %>%
    mutate(color = pal_fun(as.numeric(predictor.zip)))

  colors.legend <- zcta_level.df %>%
    arrange(predictor.zip) %>%
    distinct(predictor.zip,color)
   

  HIV_geomap = zcta_level.df %>%
    #added this line of code to prevent the following error:
    # Warning message:
    # sf layer has inconsistent datum (+proj=longlat +datum=NAD83 +no_defs).
  # Need '+proj=longlat +datum=WGS84'
    # https://community.rstudio.com/t/how-to-map-tidycensus-list-output/13547/4
    sf::st_transform(4326) %>%
    leaflet() %>%
    addPolygons(
      stroke = T, weight = 1,opacity = 0.5, color = 'white',
      fillColor = ~color,
      fillOpacity = 0.8, smoothFactor = 0.5,
      popup = p_popup) %>%
    addTiles()  %>%
    addLegend("bottomright",
              # colors = brewer.pal(break_num, "YlOrRd"),
              colors = colors.legend %>% filter(!is.na(predictor.zip)) %>% pull(color),
              labels = c(paste0('Low ',legend.name), paste0('High ',legend.name)), 
              title =  title_a) %>%
    addCircleMarkers(zcta_level.df$lng, zcta_level.df$lat, radius=zcta_level.df$n_tot/20,
                     color = "blue",
                     stroke = F, fillOpacity = 0.5)

   return(list(HIV_geomap = HIV_geomap,
              zcta_level.df = zcta_level.df))

}

zip.poverty.map <- geo_map.zipvar(county.df = county.df,
                                  county_zip_list = county_zip_list,
                                  zip_var = 'predictor_zc_per_poverty_cur_binary',
                                  'cur_zip_cd',
                                  title_a = 'High vs low poverty',
                                  legend.name = 'poverty',
                                  min.n.obs = 10)
zip.poverty.map$HIV_geomap
###plot high vs low outcomes
# outcome <- map_outcome_list[[2]]
# title_a = outcome$outcome_var_list_nice
title_a = 'High vs low proportion in-care'
county.df = county.df
county_zip_list = county_zip_list
outcome_var = 'outcome_care'
geo_zip_var = 'cur_zip_cd'
min.n.obs = 10
legend.name = 'in-care'

geo_map.highlow <- function(county.df,
                    county_zip_list,
                    outcome_var,
                    geo_zip_var,
                    title_a,
                    legend.name = 'poverty',
                    min.n.obs = 10) {

  zcta_data <- get_acs(
    geography = "zcta",
    variables = "B19013_001",
    zcta = county_zip_list,
    geometry = TRUE
  )

  
  #for outcomes such as in-care/supp/late stage dx
  county_summary.df = county.df %>%
    group_by(!! rlang::sym(geo_zip_var)) %>%
    summarize(n_tot = n(),
              n_outcome = sum(!! rlang::sym(outcome_var))) %>%
    ungroup() %>%
    filter(n_tot > min.n.obs) %>%
    mutate(per_outcome = n_outcome/n_tot) %>%
    mutate(rsd_zip_cd_cat = as.character(!! rlang::sym(geo_zip_var)),
           median.outcome = median(per_outcome),
           highlow.outcome = as.numeric(per_outcome > median.outcome))


  zcta_level.df = left_join(zcta_data,
                            county_summary.df,
                            by = c("GEOID" = "rsd_zip_cd_cat"))

  zcta_level.df = left_join(zcta_level.df,
                            zip_code_db %>% select(zipcode, lat, lng),
                            by = c("GEOID" = "zipcode")) %>%
    filter(!is.na(lat))

  # zcta_level.df.filt <- zcta_level.df %>%
  #   mutate()

  # "#808080" NA color (grey)


  break_num = 2
  pal_fun <- colorFactor("YlOrRd", NULL) #, na.color = "#808080"
  p_popup <- paste0("<strong>", title_a, ": </strong>", zcta_level.df$highlow.outcome, "<br>",
                    "<strong>Zip code: </strong>", zcta_level.df$GEOID)
  # p_popup <- paste0("<strong>", title_a, ": </strong>", zcta_level.df$per_outcome, "<br>",
  #                   "<strong>Zip code: </strong>", zcta_level.df$GEOID)
  
  zcta_level.df = zcta_level.df %>%
    mutate(color = pal_fun(as.numeric(highlow.outcome)))

  colors.legend <- zcta_level.df %>%
    arrange(highlow.outcome) %>%
    distinct(highlow.outcome,color)
   
  HIV_geomap = zcta_level.df %>%
    #added this line of code to prevent the following error:
    # Warning message:
    # sf layer has inconsistent datum (+proj=longlat +datum=NAD83 +no_defs).
  # Need '+proj=longlat +datum=WGS84'
    # https://community.rstudio.com/t/how-to-map-tidycensus-list-output/13547/4
    sf::st_transform(4326) %>%
    leaflet() %>%
    addPolygons(
      stroke = T, weight = 1,opacity = 0.5, color = 'white',
      fillColor = ~color,
      fillOpacity = 0.8, smoothFactor = 0.5,
      popup = p_popup) %>%
    addTiles()  %>%
    addLegend("bottomright",
              # colors = brewer.pal(break_num, "YlOrRd"),
              colors = colors.legend %>% filter(!is.na(highlow.outcome)) %>% pull(color),
              labels = c(paste0('Low ',legend.name), paste0('High ',legend.name)), 
              title =  title_a) %>%
    addCircleMarkers(zcta_level.df$lng, zcta_level.df$lat, radius=zcta_level.df$n_tot/20,
                     color = "blue",
                     stroke = FALSE, fillOpacity = 0.5)

  return(list(HIV_geomap = HIV_geomap,
              zcta_level.df = zcta_level.df))

}

title_a = 'High vs low proportion in-care'
county.df = county.df
county_zip_list = county_zip_list
outcome_var = 'outcome_care'
geo_zip_var = 'cur_zip_cd'
min.n.obs = 10
legend.name = 'in-care'

care.highlow.map <- geo_map.highlow(county.df = county.df,
                                    county_zip_list = county_zip_list,
                                    outcome_var = 'outcome_care',
                                    geo_zip_var = 'cur_zip_cd',
                                    title_a = 'High vs low proportion in-care',
                                    legend.name = 'in-care',
                                    min.n.obs = 10) 

supp.highlow.map <- geo_map.highlow(county.df = county.df,
                                    county_zip_list = county_zip_list,
                                    outcome_var = 'outcome_vl_supp',
                                    geo_zip_var = 'cur_zip_cd',
                                    title_a = 'High vs low proportion viral suppression',
                                    legend.name = 'rates of viral suppression',
                                    min.n.obs = 10) 
supp.highlow.map$HIV_geomap



outcome1 = zip.poverty.map
outcome2 = care.highlow.map
intx.recodings = c("Low poverty, low in-care" = "00",
             "Low poverty, high in-care" = "01",
             "High poverty, low in-care" = "10",
             "High poverty, High in-care" = "11")
title_a = 'Poverty and in-care interaction'
geo_map.intx <- function(outcome1,
                    outcome2,
                    intx.recodings,
                    title_a) {
  
  intx.df <- left_join(outcome1$zcta_level.df,outcome2$zcta_level.df %>% select(GEOID,highlow.outcome) %>% st_drop_geometry %>% as.data.frame(), by = 'GEOID')

  intx.df.fmt <- intx.df %>%
  mutate(zip.outcome.intx = paste0(predictor.zip,highlow.outcome) %>% factor(levels = c('00','01','10','11')) %>%
  fct_recode(!!!intx.recodings))
  
   break_num = 2
  pal_fun <- colorFactor("YlOrRd", NULL) #, na.color = "#808080"
  p_popup <- paste0("<strong>", title_a, ": </strong>", intx.df.fmt$zip.outcome.intx, "<br>",
                    "<strong>Zip code: </strong>", intx.df.fmt$GEOID)
  # p_popup <- paste0("<strong>", title_a, ": </strong>", zcta_level.df$per_outcome, "<br>",
  #                   "<strong>Zip code: </strong>", zcta_level.df$GEOID)
  
  intx.df.fmt = intx.df.fmt %>%
    mutate(color = pal_fun(as.numeric(zip.outcome.intx)))

  colors.legend <- intx.df.fmt %>%
    arrange(zip.outcome.intx) %>%
    distinct(zip.outcome.intx,color)
  
  HIV_geomap = intx.df.fmt %>%
    #added this line of code to prevent the following error:
    # Warning message:
    # sf layer has inconsistent datum (+proj=longlat +datum=NAD83 +no_defs).
  # Need '+proj=longlat +datum=WGS84'
    # https://community.rstudio.com/t/how-to-map-tidycensus-list-output/13547/4
    sf::st_transform(4326) %>%
    leaflet() %>%
    addPolygons(
      stroke = T, weight = 1,opacity = 0.5, color = 'white',
      fillColor = ~color,
      fillOpacity = 0.8, smoothFactor = 0.5,
      popup = p_popup) %>%
    addTiles()  %>%
    addLegend("bottomright",
              # colors = brewer.pal(break_num, "YlOrRd"),
              colors = colors.legend %>% filter(!is.na(zip.outcome.intx)) %>% pull(color),
              labels = intx.recodings %>% names() , #%>% str_replace_all('IN-CARE', 'IN-CARE<br>') %>% str_replace_all(',', '<br>')
              title =  title_a) %>%
    addCircleMarkers(intx.df.fmt$lng, intx.df.fmt$lat, radius=intx.df.fmt$n_tot/20,
                     color = "blue",
                     stroke = FALSE, fillOpacity = 0.5)
return(HIV_geomap)

}


poverty.care.map <- geo_map.intx(outcome1 = zip.poverty.map,
                                 outcome2 = care.highlow.map,
                                 intx.recodings = c("Low poverty, low in-care" = "00",
                                                    "Low poverty, high in-care" = "01",
                                                    "High poverty, low in-care" = "10",
                                                    "High poverty, High in-care" = "11"),
                                 title_a = 'Poverty and in-care interaction')
poverty.care.map
poverty.supp.map <- geo_map.intx(outcome1 = zip.poverty.map,
                                 outcome2 = supp.highlow.map,
                                 intx.recodings = c("Low poverty, low suppression" = "00",
                                                    "Low poverty, high suppression" = "01",
                                                    "High poverty, low suppression" = "10",
                                                    "High poverty, High suppression" = "11"),
                                 title_a = 'Poverty and viral suppression interaction')
poverty.supp.map
mapshot(poverty.supp.map, file = paste0(result_file_loc, 'poverty_supp_',"_map.png", sep = ""))
################
## Geographic maps
################
map_outcome_list <- list(
  "late stage dx" = list(
    outcome_var_list = c("outcome_diagnosis"),
    outcome_var_list_nice = c("Proportion Late stage dx"),
    random_effect_var_list = c("rsd_zip_cd")
  ),
  "in-care" = list(
    outcome_var_list = c("outcome_care"),
    outcome_var_list_nice = c("Proportion in-care"),
    random_effect_var_list = c("cur_zip_cd")
  ),
  "supp" = list(
    outcome_var_list = c("outcome_vl_supp"),
    outcome_var_list_nice = c("Proportion VL Suppression"),
    random_effect_var_list = c("cur_zip_cd")
  )
)

# outcome <- map_outcome_list[[1]]
map_plots <- lapply(X = map_outcome_list,FUN = function(outcome){
  # outcome_var_i = outcome_var_list[i]

  HIV_geomap = geo_map(title_a = outcome$outcome_var_list_nice,
                       county.df = county.df,
                       county_zip_list = county_zip_list,
                       outcome_var = outcome$outcome_var_list,
                       geo_zip_var = outcome$random_effect_var_list)

  mapshot(HIV_geomap, file = paste0(result_file_loc, outcome$outcome_var_list, "_map.png", sep = ""))
  return(HIV_geomap)
})
```

## Late stage diagnosis

```{r,fig.width=10,fig.height=11}
#| echo: false
#| label: late_stage_dx_map
#| fig-cap: "Heatmap of Late stage diagnosis"
map_plots$`late stage dx`
```

## In-care

```{r,fig.width=10,fig.height=11}
#| echo: false
#| label: in-care
#| fig-cap: "Heatmap of proportion in-care in 2022"
map_plots$`in-care`
```

## Viral suppression

```{r,fig.width=10,fig.height=11}
#| echo: false
#| label: viral-suppression
#| fig-cap: "Heatmap of proportion virally suppressed in 2022"
map_plots$supp
```

# Stage of care

```{r}
#| label: stage-of-care
#| echo: false
#| include: false

#load('stage.output.2023.RData')

# load('~/Documents/Projects/Nevada/data requests/NV_Data_Processing_Scripts/NV_stage_of_care_code.R')
#  ended up calcualting it elsewhere and just loading the results.
# stage.output.2023 <- stage.care(demographics = demographics,
#                                 year.stage = 2023,
#                                 county.stage = 'Clark County',
#                                 state.stage = 'NV')
# 
# 
#county.df.stage <- left_join(county.df,stage.output.2023,by = c('UCI' = 'UCI'))
# table(county.df.stage$outcome_vl_supp,county.df.stage$Stage)
# 
# table(county.df.stage$outcome_care,county.df.stage$Stage,useNA = 'always')

# county.df.stage.check.vl <- county.df.stage %>%
#   filter(VL_supp_2022_COC %in% 1 & Stage != 5)
# 
# county.df.stage.check.care <- county.df.stage %>%
#   filter((In_care_2022_COC %in% 0 & Stage %in% c(4,5)) | (In_care_2022_COC %in% 1 & !(Stage %in% c(4,5))))

# stage.tbl <- county.df.stage %>% 
#     select(Race.factor, `Birth Sex.factor`, `Exposure Category`,
#            clustered_2022,Sequence_ever,
#            Stage) %>% # keep only columns of interest
#     tbl_summary(     
#       by = Stage,                                               # stratify entire table by outcome
#       statistic = list(all_continuous() ~ "{mean} ({sd})",        # stats and format for continuous columns
#                        all_categorical() ~ "{n} / {N} ({p}%)"),   # stats and format for categorical columns
#       digits = all_continuous() ~ 1,                              # rounding for continuous columns
#       type   = all_categorical() ~ "categorical",                 # force all categorical levels to display
#       label  = list(                                              # display labels for column names
#         Race.factor   ~ "Race",                           
#         `Birth Sex.factor` ~ "Sex",
#         `Exposure Category` ~ "Transmission Risk",
#         # Education ~ "Education",
#         Sequence_ever ~ 'Sequence available',
#         clustered_2022 ~ "Genetic cluster"),
#         # outcome_diagnosis  ~ "Late-stage Diagnosis",
#         # outcome_care  ~ "In Care",
#         # outcome_vl_supp  ~ "Viral Suppression"),
#       missing_text = "Missing"                                    # how missing values should display
#     ) %>%
#   add_overall()
```

```{r}
#| echo: false
#| output: true
#| tbl-cap: Stage of care at Jan 1. 2023
#stage.tbl
```
