---
title: "HIVcountry_main"
author: "Ravi"
date: "2023-07-21"
output: html_document
---

```{r library, echo=FALSE, message=FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library('MASS')
library('tidyverse')
library('readxl')
library('kableExtra')
library('eeptools')
library('geepack')
library('caret')
library('tidycensus')
library('zipcodeR')
library('leaflet')
library('classInt')
library('RColorBrewer')
library('gtsummary')
library('lme4')
library('glmmLasso')
library('nlme')

```



```{r data, include=FALSE}

source("/Users/ravigoyal/Dropbox/Academic/Research/Projects/HIVcounty/HIVcounty/HIVcounty_func.R")

directory_loc = "/Users/ravigoyal/Dropbox/Academic/Research/Projects/HIVcounty/"

county_demo_file_subloc = "Clark_County_data/clark_county_demographics.csv"
county_zip_file_subloc = "Clark_County_data/clark_county_by_zip.csv"
county_vl_file_subloc = "Clark_County_data/clark_viral_suppression_by_year.csv"

county_zip_list = search_county("Clark", "NV") %>% pull(zipcode)

#county_demo_file_subloc = "San_Diego_data/sd_county_demographics.csv"
#county_zip_file_subloc = "San_Diego_data/sd_county_by_zip.csv"
#county_vl_file_subloc = "San_Diego_data/sd_viral_suppression_by_year.csv"

outcome_var = "HIVstage"
outcome_var = "VLsupp"
title_a = "Proportion VL Suppression"
lab_x_a = "Proportion VL Suppression"
lab_y_a = "Proportion VL Suppression"
```

## Create analytic dataset

We create an analytic dataset that includes individual and zip code level covariates. 

```{r analytic_file, echo=FALSE, message=FALSE, warning = FALSE}

county.df  = read_county_data(directory_loc = directory_loc,
                             county_demo_file_subloc = county_demo_file_subloc,
                             county_zip_file_subloc = county_zip_file_subloc,
                             county_zip_list = county_zip_list,
                             county_vl_file_subloc = county_vl_file_subloc
                             ) 
```     

Number of individuals after filtering to diagnosed in county: `r county_diagnosis.df %>% nrow()`

## Research question

Research Question 3: Identify predictors (including viral clustering) that are associated with high rates of stage 3-4 new diagnoses. 

## Outcome generation

We dichotomize the HIV stage of an individual.

* HIV ONLY (HIV, STAGE 1 OR 2) 

VS

* HIV AND AIDS SIMULTANEOUSLY (HIV, STAGE 3)
* HIV AND LATER AIDS (HIV, STAGE 3)

```{r outcomes_diagnosis, echo=FALSE, message=FALSE, warning = FALSE}

county.df = county.df %>%
  mutate(outcome_diagnosis = case_when(
    `Disease Category` == "HIV ONLY (HIV, STAGE 1 OR 2)" ~ 0,
    `Disease Category` == "HIV AND AIDS SIMULTANEOUSLY (HIV, STAGE 3)" ~ 1,
    `Disease Category` == "HIV AND LATER AIDS (HIV, STAGE 3)" ~ 1,
    TRUE ~ NA_real_))

county.df = county.df %>%
  mutate(outcome_care = as.numeric(`2022` %in% c(0,1))) 

county.df = county.df %>%
  mutate(outcome_TEMP = `2022`) %>%
  mutate(outcome_vl_supp = case_when(
    outcome_TEMP == 0 ~ 0,
    outcome_TEMP == 1 ~ 1,
    TRUE ~ 0))

```

## Descriptive statistics

The number of individuals in Stages 1 or 2 (outcome = 0) vs. number of individuals in Stage 3

```{r desc_stats, echo=FALSE, message=FALSE, warning = FALSE}

county.df %>% 
  select(Race.factor, `Birth Sex.factor`, `Exposure Category`, Education,
         clustered_2022,
         outcome_diagnosis, outcome_care, outcome_vl_supp) %>% # keep only columns of interest
  tbl_summary(     
    #by = outcome,                                               # stratify entire table by outcome
    statistic = list(all_continuous() ~ "{mean} ({sd})",        # stats and format for continuous columns
                     all_categorical() ~ "{n} / {N} ({p}%)"),   # stats and format for categorical columns
    digits = all_continuous() ~ 1,                              # rounding for continuous columns
    type   = all_categorical() ~ "categorical",                 # force all categorical levels to display
    label  = list(                                              # display labels for column names
      Race.factor   ~ "Race",                           
      `Birth Sex.factor` ~ "Sex",
      `Exposure Category` ~ "Transmission Risk",
      Education ~ "Education",
      clustered_2022 ~ "Genetic cluster",
      outcome_diagnosis  ~ "Late-stage Diagnosis",
      outcome_care  ~ "In Care",
      outcome_vl_supp  ~ "Viral Suppression"),
    missing_text = "Missing"                                    # how missing values should display
  )
  
```

### Predictors

```{r predictors, echo=FALSE, message=FALSE, warning = FALSE}

county.df = county.df %>%
  mutate(predictor_education = case_when(
    Education == "<= 8TH GRADE" ~ 0,
    Education == "SOME SCHOOL, LEVEL UNKNOWN" ~ 0,
    Education == "SOME HIGH SCHOOL" ~ 0,
    Education == "HIGH SCHOOL GRAD" ~ 0,
    Education == "SOME COLLEGE" ~ 1,
    Education == "COLLEGE DEGREE" ~ 1,
    Education == "POST-GRADUATE WORK" ~ 1,
    TRUE ~ NA_real_))

county.df = county.df %>%
  mutate(predictor_hispanic = case_when(
    Race == "WHITE" ~ 0,
    Race == "OTHER" ~ 0,
    Race == "BLACK" ~ 0,
    Race == "HISPANIC, ANY RACE" ~ 1,
    TRUE ~ NA_real_))

county.df = county.df %>%
  mutate(predictor_msm = case_when(
    `Exposure Category` == "MSM" ~ 1,
    `Exposure Category` == "IDU" ~ 0,
    `Exposure Category` == "NO REPORTED RISK" ~ 0,
    `Exposure Category` == "MSM & IDU" ~ 1,
    `Exposure Category` == "HETEROSEXUAL CONTACT" ~ 0,
    `Exposure Category` == "OTHER" ~ 0,
    `Exposure Category` == "PERINATAL EXPOSURE" ~ 0,
    TRUE ~ NA_real_))

county.df = county.df %>%
  mutate(predictor_idu = case_when(
    `Exposure Category` == "MSM" ~ 0,
    `Exposure Category` == "IDU" ~ 1,
    `Exposure Category` == "NO REPORTED RISK" ~ 0,
    `Exposure Category` == "MSM & IDU" ~ 1,
    `Exposure Category` == "HETEROSEXUAL CONTACT" ~ 0,
    `Exposure Category` == "OTHER" ~ 0,
    `Exposure Category` == "PERINATAL EXPOSURE" ~ 0,
    TRUE ~ NA_real_))

county.df = county.df %>%
  mutate(predictor_diagnosis_year = `Diagnosis year`)

county.df = county.df %>%
  mutate(predictor_birth_sex = case_when(
    `Birth Sex` == "FEMALE" ~ 0,
    `Birth Sex` == "MALE" ~ 1,
    TRUE ~ NA_real_))

```

## Individual-level analyses

### Regression

Individual-level variables to consider: 
* predictor_diagnosis_year
* Age.diagnosis
* predictor_hispanic
* predictor_idu
* predictor_msm
* predictor_education
* predictor_birth_sex
* clustered_2022

zip code-level variables to consider: 
* Median income
* Percent in poverty
* Percent clustered in 2022

First set of estimates and p-values are for univariate regression. The second set is for multivariate regression.

```{r Univariate_reg, echo=FALSE, message=FALSE, warning = FALSE}

uni_reg_res.df = tibble(
  variable = NULL,
  estimate_uni = NULL,
  p_val_uni = NULL
)

reg_variable_ind.vec = c(
                 "predictor_diagnosis_year",
                 "Age.diagnosis",
                 "predictor_hispanic",
                 "predictor_idu",
                 "predictor_msm",
                 "predictor_education",
                 "predictor_birth_sex",
                 "clustered_2022")

reg_variable_zip.vec = c(
                 "Median_Income_Total",
                 "predictor_county_per_poverty",
                 "percent_clustered_2022"
                 )

reg_variable.vec = c(reg_variable_ind.vec, reg_variable_zip.vec)

for (reg_variable_ind in reg_variable.vec) {

  regression_formula <- paste0("outcome_diagnosis ~ ", reg_variable_ind, " + (1 | rsd_zip_cd)")

  uni_reg = glmer(as.formula(regression_formula), 
              family = binomial,
              data = county.df)

  uni_reg_sum = summary(uni_reg)
  uni_reg_res_TEMP.df = tibble(
    variable = reg_variable_ind,
    estimate_uni = uni_reg_sum$coefficients[2,c(1)],
    p_val_uni = uni_reg_sum$coefficients[2,c(4)]
  )
  
  uni_reg_res.df = bind_rows(uni_reg_res.df, uni_reg_res_TEMP.df)
}

```


```{r multivariate_reg, echo=FALSE, message=FALSE, warning = FALSE}

county.df_TEMP = county.df %>% 
  select(reg_variable.vec, outcome_diagnosis, rsd_zip_cd) %>%
  mutate(rsd_zip_cd_fac = as.factor(rsd_zip_cd)) %>%
  na.omit()

regression_formula <- as.formula(paste0("outcome_diagnosis", " ~ ", 
                                        paste0(reg_variable.vec, 
                                               collapse = " + ")))

N<-nrow(county.df_TEMP)
ind<-sample(N,N)
lambda <- seq(500,0,by=-5)

family <- binomial(link = logit)

kk<-5
nk <- floor(N/kk)

Devianz_ma<-matrix(Inf,ncol=kk,nrow=length(lambda))

## first fit good starting model
PQL<-glmmPQL(outcome_diagnosis~1,
               random = ~1|rsd_zip_cd_fac,
               family=family,
               data=county.df_TEMP)

Delta.start<-c(as.numeric(PQL$coef$fixed),rep(0,6),as.numeric(t(PQL$coef$random$rsd_zip_cd_fac)))
Q.start<-as.numeric(VarCorr(PQL)[1,1])

## loop over the folds  
for(j in 1:length(lambda))
{
  print(paste("Iteration ", j,sep=""))
  
  for (i in 1:kk)
  {
    if (i < kk)
    {
      indi <- ind[(i-1)*nk+(1:nk)]
    }else{
      indi <- ind[((i-1)*nk+1):N]
    }
    
    county.train<-county.df_TEMP[-indi,]
    county.test<-county.df_TEMP[indi,]
    
    glm2 <- try(glmmLasso(fix = regression_formula,
                          rnd = list(rsd_zip_cd_fac=~1),
                          family = binomial(link = logit), 
                          data =county.df_TEMP,
                          lambda=lambda[j],
                          switch.NR=FALSE,
                          final.re=FALSE,
                          control=list()), #list(start=Delta.start,q_start=Q.start)),
                silent=TRUE) 
    
    if(!inherits(glm2, "try-error"))
    {  
      y.hat<-predict(glm2,as.data.frame(county.test))     
      Devianz_ma[j,i]<-sum(family$dev.resids(county.test$outcome_diagnosis,y.hat,wt=rep(1,length(y.hat))))
    }
  }
  print(sum(Devianz_ma[j,]))
}

Devianz_vec<-apply(Devianz_ma,1,sum)
opt2<-which.min(Devianz_vec)
       
glm2_final <- glmmLasso(fix = regression_formula,
                  rnd = list(rsd_zip_cd_fac=~1),
                  family = binomial(link = logit), 
                  data =county.df_TEMP,
                  lambda=lambda[opt2],
                  switch.NR=FALSE,
                  final.re=TRUE,
                  control=list()) #list(start=Delta.start,q_start=Q.start))

multi_reg_sum = summary(glm2_final)

multi_reg_res.df = tibble(
  variable = rownames(multi_reg_sum$coefficients)[c(2:nrow(multi_reg_sum$coefficients))],
  estimate_multi = multi_reg_sum$coefficients[2:nrow(multi_reg_sum$coefficients),c(1)],
  p_val_multi = multi_reg_sum$coefficients[2:nrow(multi_reg_sum$coefficients),c(4)]
)

full_join(uni_reg_res.df, multi_reg_res.df, by = "variable") %>%
  kbl() %>%
  kable_paper("hover", full_width = F)

```

### Geographic map

```{r desc_map, echo=FALSE, message=FALSE, warning = FALSE}

zcta_data <- get_acs(
  geography = "zcta",
  variables = "B19013_001",
  zcta = county_zip_list, 
  geometry = TRUE
)

county_summary.df = county.df %>%
  group_by(rsd_zip_cd) %>%
  summarize(n_tot = n(),
            n_outcome = sum(outcome)) %>%
  ungroup() %>%
  mutate(per_outcome = n_outcome/n_tot) %>%
  mutate(rsd_zip_cd_cat = as.character(rsd_zip_cd))
  

zcta_level.df = left_join(zcta_data,
                          county_summary.df,
                          by = c("GEOID" = "rsd_zip_cd_cat"))

zcta_level.df = left_join(zcta_level.df,
                          zip_code_db %>% select(zipcode, lat, lng),
                          by = c("GEOID" = "zipcode")) %>%
  filter(!is.na(lat))

break_num = 3
pal_fun <- colorQuantile("YlOrRd", NULL, n = break_num)
p_popup <- paste0("<strong>", title_a, ": </strong>", zcta_level.df$per_outcome)
breaks_qt <- classIntervals(zcta_level.df$per_outcome, n = break_num, style = "quantile")

leaflet(zcta_level.df) %>%
  addPolygons(
    stroke = FALSE, 
    fillColor = ~pal_fun(per_outcome),
    fillOpacity = 0.8, smoothFactor = 0.5,
    popup = p_popup) %>%
  addTiles()  %>%
  addLegend("bottomright", 
            colors = brewer.pal(break_num, "YlOrRd"), 
            labels = paste0("up to ", format(breaks_qt$brks[-1], digits = 2)),
            title =  title_a) %>%
  addCircleMarkers(zcta_level.df$lng, zcta_level.df$lat, radius=zcta_level.df$n_tot/50, 
                   color = "blue", 
                   stroke = FALSE, fillOpacity = 0.5) 

```

### Exploration


```{r desc_demo, echo=FALSE, message=FALSE, warning = FALSE}

race_prop.df = county_diagnosis.df %>%
  group_by(Race.full.factor) %>%
  summarize(total_n = n(),
            late_diag_n = sum(outcome)) %>%
  ungroup() %>%
  mutate(prop = late_diag_n/total_n)

p_race <-ggplot(race_prop.df, aes(x=Race.full.factor, y=prop, fill=Race.full.factor)) +
  geom_bar(stat="identity")+theme_minimal() +
  geom_text(aes(label=total_n), vjust=-1) +
  labs(title="",
       x="Race / Ethnicity", y = "Proportion Late Diagnosis") +
  ylim(0,1) +
  theme(axis.title.x = element_text(size = 16),
        axis.text.x = element_blank(),
        axis.title.y = element_text(size = 16)) +
  guides(fill=guide_legend(title="Race / Ethnicity"))+
  scale_fill_brewer(palette="Spectral")
p_race

exp_cat_prop.df = county_diagnosis.df %>%
  group_by(`Exposure Category.factor`) %>%
  summarize(total_n = n(),
            late_diag_n = sum(outcome)) %>%
  ungroup() %>%
  mutate(prop = late_diag_n/total_n)

p_exp_cat <-ggplot(exp_cat_prop.df, aes(x=`Exposure Category.factor`, y=prop, fill=`Exposure Category.factor`)) +
  geom_bar(stat="identity")+theme_minimal() +
  geom_text(aes(label=total_n), vjust=-1) +
  labs(title="",
       x="Exposure Category", y = lab_y_a) +
  ylim(0,1) +
  theme(axis.title.x = element_text(size = 16),
        axis.text.x = element_blank(),
        axis.title.y = element_text(size = 16)) +
  guides(fill=guide_legend(title="Exposure Risk")) +
  scale_fill_brewer(palette="Spectral")
p_exp_cat


edu_cat_prop.df = county_diagnosis.df %>%
  group_by(Education.factor) %>%
  summarize(total_n = n(),
            late_diag_n = sum(outcome)) %>%
  ungroup() %>%
  mutate(prop = late_diag_n/total_n)

p_edu_cat <-ggplot(edu_cat_prop.df, aes(x=Education.factor, y=prop, fill=Education.factor)) +
  geom_bar(stat="identity")+theme_minimal() +
  geom_text(aes(label=total_n), vjust=-1) +
  labs(title="",
       x="Education Category", y = lab_y_a) +
  ylim(0,1) +
  theme(axis.title.x = element_text(size = 16),
        axis.text.x = element_blank(),
        axis.title.y = element_text(size = 16)) +
  guides(fill=guide_legend(title="Education")) +
  scale_fill_brewer(palette="Spectral")
p_edu_cat


```


```{r exploratory_analysis, echo=FALSE, message=FALSE, warning = FALSE}

p_predictor_diagnosis_year <- ggplot(county_diagnosis.df, aes(x=predictor_diagnosis_year, y=outcome)) +
  geom_point(col = "blue") +
  geom_smooth(col = "red") +
  labs(title=paste0(lab_y_a, "by Diagnosis Year"),
        x ="Diagnosis Year", y = lab_y_a)

p_predictor_diagnosis_year

p_Age.diagnosis <- ggplot(county_diagnosis.df, aes(x=Age.diagnosis, y=outcome)) +
  geom_point() +
  geom_smooth() +
  labs(title=paste0(lab_y_a, "by Age at Diagnosis"),
        x ="Age at Diagnosis", y = lab_y_a)

p_Age.diagnosis

p_zip_income <- ggplot(county_diagnosis.df, aes(x=Median_Income_Total, y=outcome)) +
  geom_point() +
  geom_smooth() +
  labs(title=paste0(lab_y_a, "by Zip Code Level Income"),
        x ="Zip Code Level Income", y = lab_y_a)

p_zip_income

p_percent_clustered_2022 <- ggplot(county_diagnosis.df, aes(x=percent_clustered_2022, y=outcome)) +
  geom_point() +
  geom_smooth() +
  labs(title=paste0(lab_y_a, "by Zip Code Clustering"),
        x ="Zip Code Level Clustering", y = lab_y_a)

p_percent_clustered_2022

county_diagnosis.df %>%
  group_by(clustered_2022, outcome) %>%
  summarize(n=n()) 

```


