---
title: "SanDiego_County_main"
output: html_document
date: "2023-05-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

directory_loc = "/Users/ravigoyal/Dropbox/Academic/Research/Projects/HIVcounty/"
county_demo_file_subloc = "San_Diego_data/sd_county_demographics.csv"
county_zip_file_subloc = "San_Diego_data/sd_county_by_zip.csv"

```


```{r library, echo=FALSE, message=FALSE, warning = FALSE}
library('tidyverse')
library('readxl')
library('kableExtra')
library('eeptools')
library('geepack')
library('caret')
library('tidycensus')
library('zipcodeR')

source("/Users/ravigoyal/Dropbox/Academic/Research/Projects/HIVcounty/HIVcounty/HIVcounty_func.R")
```

## Research question

Research Question 3: Identify predictors (including viral clustering) that are associated with high rates of stage 3-4 new diagnoses. 

## Create analytic dataset

We create an analytic dataset that includes individual and zip code level covariates. 

```{r dataset, echo=FALSE, message=FALSE, warning = FALSE}

county_zip_list = search_county("San Diego", "CA") %>% pull(zipcode)

county_diagnosis.df  = read_county_data(directory_loc = directory_loc,
                             county_demo_file_subloc = county_demo_file_subloc,
                             county_zip_file_subloc = county_zip_file_subloc,
                             county_zip_list = county_zip_list
                             ) 

```     

Number of individuals after filtering to diagnosed in county: `r county_diagnosis.df %>% nrow()`

## Outcome generation

We create a binary variable indicating if an individual was sequenced within 30-days of diagnosis.

```{r outcomes, echo=FALSE, message=FALSE, warning = FALSE}

```

